---
title: "Create Long Format Panel Dataset"
author: "Migration Networks UCNets Project"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

# Overview

This notebook creates the LONG format panel dataset for the Migration Networks analysis.
It loads raw panel data and sources three processing scripts:

1. **01_create_long_variables.Rmd** - Creates lagged, cumulative, and transition variables
2. **02_create_long_decomposition.Rmd** - Creates between/within decomposition
3. **03_merge_external_data.Rmd** - Merges external datasets

## Load Libraries

```{r libraries}
library(haven)
library(dplyr)
library(tidyr)
library(here)
```

## Load Raw Data

```{r load-data}
cat("Loading raw data...\n")
df <- read_dta(here("data", "R_LONG_Moving_Paper_CS_file_24_02_24.dta"))

# Ensure data is sorted by prim_key and WAVE
df <- df %>% arrange(prim_key, WAVE)

cat(sprintf("Loaded %d observations for %d unique individuals\n",
            nrow(df),
            n_distinct(df$prim_key)))
```

## Create Variables

Sources lagged variables, cumulative variables, transition variables,
derived variables, and variable aliases.

```{r create-variables, child=here('code', '01_create_long_variables.Rmd')}
```

## Panel Decomposition

Sources between/within decomposition, demographics, weights, and data saving.

```{r panel-decomposition, child=here('code', '02_create_long_decomposition.Rmd')}
```

## Merge External Data

Sources external datasets merged on prim_key_wave.

```{r merge-external, child=here('code', '03_merge_external_data.Rmd')}
```

## Save Final Dataset

```{r save-data}
# Save the complete dataset (after all processing and merges)
saveRDS(df, here("data", "out", "R_DF.rds"))
cat("\nDataset saved to data/out/R_DF.rds\n")
```

## Summary

```{r summary}
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("DATA PROCESSING COMPLETE\n")
cat(paste(rep("=", 70), collapse = ""), "\n")

cat(sprintf("\nFinal dataset: %d observations, %d variables\n",
            nrow(df), ncol(df)))

cat("\nKey outcome variables:\n")
cat("  - Moved_w23: Any move between waves 2-3\n")
cat("  - Moved_in_Bay_w23: Moved within Bay Area\n")
cat("  - Moved_out_of_Bay_w23: Moved out of Bay Area\n")

cat("\nData file saved:\n")
cat("  - data/out/R_DF.rds\n")
```
