---
title: "Create Panel Decomposition"
author: "Migration Networks UCNets Project"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup-decomp, include=FALSE}
# This file is designed to be sourced as a child document from 00_create_long_data.Rmd
# It expects 'df' to already be loaded in the environment
```

# Panel Decomposition

This notebook creates between-person means and within-person deviations for panel analysis.
Also expands time-invariant demographics and creates Wave 3 weights.

## Load Helper Functions

```{r load-helpers-decomp}
library(here)
source(here("code", "helpers", "panel_helpers.R"))
```

## Between/Within Person Decomposition

Creates means (between) and deviations from mean (within) for each variable.

```{r define-decomposition-vars}
vars_for_decomposition <- c(
  # Current variables
  "LE_MovedTown", "Owns_home", "Full_Time", "LT_BA", "BA_or_greater", "LE_NewJob", "LE_NewSchool",
  "LEX_Birth", "LE_FinanProb", "Marr_Cohab", "Kin_within_60_N", "Nonkin_within_60_N",
  "Nonkin_in_HH_N", "Kin_in_HH_N", "Kin_GT_Hour_N", "Nonkin_GT_Hour_N",
  "Kin_GT_Hour", "Nonkin_GT_Hour",  # Aliases without _N
  "Mother_in_Hour", "Father_in_Hour", "Plan_to_Move", "web",

  # Lagged variables
  "lag_Moved_Town", "lag_Owns_Home", "lag_Full_Time", "lag_LT_BA", "lag_BA_or_greater", "lag_LEX_Birth",
  "lag_Finan_Prob", "lag_Marr_Cohab", "lag_Kin_within_60_N", "lag_Nonkin_within_60_N",
  "lag_Nonkin_in_HH_N", "lag_Kin_in_HH_N",
  "lag_Kin_GT_Hour", "lag_Nonkin_GT_Hour",  # Aliases without _N
  "lag_Kin_GT_Hour_N", "lag_Nonkin_GT_Hour_N",  # Original names with _N
  "lag_Plan_Move",

  # Transition variables
  "Homeowner_to_renter", "Renter_to_homeowner", "Fulltime_to_less",
  "Married_to_divorced", "lag_Married_to_divorced", "lag_Fulltime_to_less",
  "lag_Homeowner_to_renter", "lag_Renter_to_homeowner", "Not_to_Married",
  "lag_Not_to_Married", "Less_to_fulltime", "lag_Less_to_fulltime",

  # Age
  "age",

  # Cumulative variables
  "cu_NewJob", "cu_NewSchool", "cu_Plan_to_Move", "cu_FinanProb", "cu_Birth",
  "lag_cu_NewJob", "lag_cu_NewSchool", "lag_cu_Plan_to_Move", "lag_cu_FinanProb",
  "lag_cu_Birth",

  # Other
  "lag_New_School", "lag_New_Job", "employed", "past_finan_prob",
  "FinanProb_to_not", "Not_to_FinanProb", "lag_FinanProb_to_not", "lag_Not_to_FinanProb"
)
```

```{r create-decomposition}
# Use helper function for batch decomposition
df <- create_between_within_multi(df, vars_for_decomposition, id_var = "prim_key")
```

## Time-Invariant Demographic Variables

Expand across all waves for each individual.

```{r time-invariant-demographics}
# Race/ethnicity variables
df <- expand_time_invariant(df, "Eth_black", "black_race")
df <- expand_time_invariant(df, "Eth_asian", "asian_race")
df <- expand_time_invariant(df, "Eth_other", "other_race")
df <- expand_time_invariant(df, "Eth_latino", "latino_eth")

# Gender (1=Male, 2=Female in original coding)
df <- df %>%
  group_by(prim_key) %>%
  mutate(gender = first(na.omit(gender))) %>%
  ungroup() %>%
  mutate(female = ifelse(gender == 2, 1, 0))
```

## Wave 3 Weights

Create Wave 3 demographic weights for all observations.

```{r wave3-weights}
df <- df %>%
  group_by(prim_key) %>%
  mutate(
    wt_dem_95_inf_w3 = max(if_else(WAVE == 3, wt_dem_95_inf, NA_real_), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(wt_dem_95_inf_w3 = if_else(is.infinite(wt_dem_95_inf_w3), NA_real_, wt_dem_95_inf_w3))

cat("\nPanel decomposition complete.\n")
```
