---
title: "Merge External Data"
author: "Migration Networks UCNets Project"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup-merge, include=FALSE}
# This file is designed to be sourced as a child document from 00_create_long_data.Rmd
# It expects 'df' to already be loaded in the environment
```

# External Data Merge

This notebook merges external datasets into the LONG format panel data.

**Merge key:** `prim_key_wave` (combination of prim_key and WAVE, no underscore separator)

## Load Helper Functions

```{r load-helpers-merge}
library(here)
library(readxl)
source(here("code", "helpers", "panel_helpers.R"))
```

## Create Merge Key in Main Data

```{r create-merge-key}
# Create prim_key_wave without underscore to match external data format
df <- df %>%
  mutate(prim_key_wave = paste0(as.character(prim_key), WAVE))

cat(sprintf("Created prim_key_wave merge key for %d observations\n", nrow(df)))
```

## External Data Source 1: Steps to Make Friends (a19f)

Open-ended responses about steps taken to meet new people, with coded categories.

```{r load-a19f}
cat("\n--- Merging Steps to Make Friends (a19f) ---\n")

a19f <- read.csv(here("data", "in", "a19f_Master_with_prim_key.csv"), stringsAsFactors = FALSE)
cat(sprintf("Loaded %d records\n", nrow(a19f)))

# Convert prim_key_wave to character for matching
a19f <- a19f %>%
  mutate(prim_key_wave = as.character(prim_key_wave))
```

### Check for Duplicates

```{r check-duplicates}
n_dups <- sum(duplicated(a19f$prim_key_wave))
if (n_dups > 0) {
  warning(sprintf("a19f has %d duplicate prim_key_wave values", n_dups))
}
```

### Rename and Select Columns

```{r rename-columns}
a19f <- a19f %>%
  rename(
    a19f_response = Response,
    a19f_religious = Religious.Community.Involvement,
    a19f_local_venue = Local.Venue.Visitation,
    a19f_volunteer = Secular.Volunteer.Service,
    a19f_online = Online.Social.Networking,
    a19f_workplace = Workplace.Socializing,
    a19f_athletic = Athletic.Engagement,
    a19f_cultural_hobby = Cultural.Hobby.Participation,
    a19f_neighborhood = Neighborhood.Integration,
    a19f_events_groups = Broader.Event.and.Group.Participation,
    a19f_through_connections = Friendship.Through.Connections,
    a19f_non_engagement = Non.Engagement.Preference,
    a19f_alternative = Alternative.Strategies,
    a19f_not_clear = Not.Clear...Uncodeable.Responses,
    a19f_no_match = No_Match
  ) %>%
  select(prim_key_wave, a19f_response,
         a19f_religious, a19f_local_venue, a19f_volunteer, a19f_online,
         a19f_workplace, a19f_athletic, a19f_cultural_hobby, a19f_neighborhood,
         a19f_events_groups, a19f_through_connections, a19f_non_engagement,
         a19f_alternative, a19f_not_clear, a19f_no_match)
```

### Check for Unmatched Records

```{r check-unmatched}
unmatched_right <- a19f %>%
  filter(!(prim_key_wave %in% df$prim_key_wave))

if (nrow(unmatched_right) > 0) {
  cat(sprintf("WARNING: %d records in a19f have no match in main data:\n", nrow(unmatched_right)))
  print(unmatched_right %>% select(prim_key_wave, a19f_response) %>% head(20))
  if (nrow(unmatched_right) > 20) {
    cat(sprintf("... and %d more\n", nrow(unmatched_right) - 20))
  }
} else {
  cat("All a19f records matched to main data.\n")
}
```

### Merge

```{r merge-a19f}
n_before <- nrow(df)
df <- df %>%
  left_join(a19f, by = "prim_key_wave")
stopifnot(nrow(df) == n_before)

n_matched <- sum(!is.na(df$a19f_response))
cat(sprintf("Matched: %d of %d main observations (%.1f%%)\n",
            n_matched, nrow(df), 100 * n_matched / nrow(df)))
```

### Create Collapsed Friend-Making Categories

Based on empirical clustering analysis (see `code/explore_a19f_categories.R`), combine
related categories into broader groupings. Creates both granular and combined versions.

**Empirical basis:**
- Hierarchical clustering groups workplace + through_connections (existing network ties)
- Activity-based strategies (athletic, cultural_hobby, events_groups, etc.) cluster together
- Non-engagement stands alone
- PCA shows engagement vs. non-engagement as primary dimension

```{r collapse-friend-categories}
df <- df %>%
  mutate(
    # -----------------------------------------------------------------
    # GRANULAR CATEGORIES (5 categories + non-engagement + unclear)
    # -----------------------------------------------------------------

    # How_Workplace: Workplace socializing (kept separate due to high frequency)
    How_Workplace = case_when(
      a19f_workplace == 1 ~ 1,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # How_Through_Connections: Meeting friends through existing connections
    How_Through_Connections = case_when(
      a19f_through_connections == 1 ~ 1,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # How_Community_Orgs: Religious + Volunteer (community-based organizations)
    How_Community_Orgs = case_when(
      a19f_religious == 1 | a19f_volunteer == 1 ~ 1,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # How_Activities: Athletic + Cultural/Hobby + Events/Groups (activity-based)
    How_Activities = case_when(
      a19f_athletic == 1 | a19f_cultural_hobby == 1 | a19f_events_groups == 1 ~ 1,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # How_Local_Places: Local Venue + Neighborhood (place-based socializing)
    How_Local_Places = case_when(
      a19f_local_venue == 1 | a19f_neighborhood == 1 ~ 1,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # How_Online: Online social networking (kept separate)
    How_Online = case_when(
      a19f_online == 1 ~ 1,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # How_Non_Engagement: Prefers not to actively seek new friends
    How_Non_Engagement = case_when(
      a19f_non_engagement == 1 ~ 1,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # How_Other_Unclear: Alternative strategies + Not clear/uncodeable
    How_Other_Unclear = case_when(
      a19f_alternative == 1 | a19f_not_clear == 1 ~ 1,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # -----------------------------------------------------------------
    # BROADER COMBINED CATEGORIES (4 categories)
    # -----------------------------------------------------------------

    # How_Network_Ties: Leveraging existing social structures
    # (Workplace + Through Connections + Religious - these cluster together)
    How_Network_Ties = case_when(
      a19f_workplace == 1 | a19f_through_connections == 1 | a19f_religious == 1 ~ 1,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # How_Active_Outreach: All proactive activity-based strategies
    # (Local venues, volunteer, online, athletic, cultural, neighborhood, events)
    How_Active_Outreach = case_when(
      a19f_local_venue == 1 | a19f_volunteer == 1 | a19f_online == 1 |
      a19f_athletic == 1 | a19f_cultural_hobby == 1 | a19f_neighborhood == 1 |
      a19f_events_groups == 1 ~ 1,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # How_Any_Engagement: Any active strategy (opposite of non-engagement)
    # Useful for simple engaged vs. not engaged analysis
    How_Any_Engagement = case_when(
      a19f_non_engagement == 0 & !is.na(a19f_response) ~ 1,
      a19f_non_engagement == 1 ~ 0,
      is.na(a19f_response) ~ NA_real_,
      TRUE ~ 0
    )
  )

cat("Created collapsed friend-making categories:\n\n")
cat("GRANULAR (7 categories):\n")
cat("  How_Workplace          - Workplace socializing\n")
cat("  How_Through_Connections - Through existing connections\n")
cat("  How_Community_Orgs     - Religious + Volunteer organizations\n")
cat("  How_Activities         - Athletic + Cultural/Hobby + Events/Groups\n")
cat("  How_Local_Places       - Local Venue + Neighborhood\n")
cat("  How_Online             - Online social networking\n")
cat("  How_Non_Engagement     - Prefers not to seek new friends\n")
cat("  How_Other_Unclear      - Alternative + Not clear\n")
cat("\nBROADER (3 categories):\n")
cat("  How_Network_Ties       - Workplace + Through Connections + Religious\n")
cat("  How_Active_Outreach    - All proactive activity-based strategies\n")
cat("  How_Any_Engagement     - Any active strategy (binary: engaged vs not)\n")
```

## External Data Source 2: Reasons for Moving (a19i)

Open-ended responses about reasons for moving, with coded categories.

```{r load-a19i}
cat("\n--- Merging Reasons for Moving (a19i) ---\n")

a19i <- read_excel(here("data", "in", "Final_Merged_Codes_a19i_alternate.xlsx"))
cat(sprintf("Loaded %d records\n", nrow(a19i)))

# Convert prim_key_wave to character for matching
a19i <- a19i %>%
  mutate(prim_key_wave = as.character(as.numeric(prim_key_wave)))
```

### Check for Duplicates (a19i)

```{r check-duplicates-a19i}
n_dups_a19i <- sum(duplicated(a19i$prim_key_wave))
if (n_dups_a19i > 0) {
  warning(sprintf("a19i has %d duplicate prim_key_wave values", n_dups_a19i))
}
```

### Rename and Select Columns (a19i)

```{r rename-columns-a19i}
# Rename by position to handle corrupted column names
new_names <- c("prim_key", "WAVE", "prim_key_wave", "a19i_response",
               "a19i_partner_spouse", "a19i_relationship_end", "a19i_own_job_school",
               "a19i_partner_job_school", "a19i_financial", "a19i_housing",
               "a19i_new_environment", "a19i_family_support", "a19i_meet_friends",
               "a19i_conflicts", "a19i_forced", "a19i_other", "a19i_not_clear",
               "a19i_forced_financial", "a19i_not_forced_financial",
               "a19i_financial_housing", "a19i_non_financial_housing",
               "a19i_financial_family_support", "a19i_non_financial_family_support")
names(a19i) <- new_names

a19i <- a19i %>%
  select(prim_key_wave, a19i_response,
         a19i_partner_spouse, a19i_relationship_end, a19i_own_job_school,
         a19i_partner_job_school, a19i_financial, a19i_housing,
         a19i_new_environment, a19i_family_support, a19i_meet_friends,
         a19i_conflicts, a19i_forced, a19i_other, a19i_not_clear,
         a19i_forced_financial, a19i_not_forced_financial,
         a19i_financial_housing, a19i_non_financial_housing,
         a19i_financial_family_support, a19i_non_financial_family_support)
```

### Check for Unmatched Records (a19i)

```{r check-unmatched-a19i}
unmatched_right_a19i <- a19i %>%
  filter(!(prim_key_wave %in% df$prim_key_wave))

if (nrow(unmatched_right_a19i) > 0) {
  cat(sprintf("WARNING: %d records in a19i have no match in main data:\n", nrow(unmatched_right_a19i)))
  print(unmatched_right_a19i %>% select(prim_key_wave, a19i_response) %>% head(20))
  if (nrow(unmatched_right_a19i) > 20) {
    cat(sprintf("... and %d more\n", nrow(unmatched_right_a19i) - 20))
  }
} else {
  cat("All a19i records matched to main data.\n")
}
```

### Merge (a19i)

```{r merge-a19i}
n_before <- nrow(df)
df <- df %>%
  left_join(a19i, by = "prim_key_wave")
stopifnot(nrow(df) == n_before)

n_matched_a19i <- sum(!is.na(df$a19i_response))
cat(sprintf("Matched: %d of %d main observations (%.1f%%)\n",
            n_matched_a19i, nrow(df), 100 * n_matched_a19i / nrow(df)))
```

### Create Collapsed Move Reason Categories

Based on suggested collapsing (4/9/24), combine related move reason categories.

```{r collapse-move-reasons}
df <- df %>%
  mutate(
    # Why_Spo_Part_&_End: Partner/Spouse OR Relationship End
    Why_Spo_Part_End = case_when(
      a19i_partner_spouse == 1 | a19i_relationship_end == 1 ~ 1,
      is.na(a19i_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # Why_Own_&_Part_Career: Own Job/School OR Partner's Job/School
    Why_Own_Part_Career = case_when(
      a19i_own_job_school == 1 | a19i_partner_job_school == 1 ~ 1,
      is.na(a19i_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # Why_Financial: Keep as is (already exists as a19i_financial)
    Why_Financial = case_when(
      a19i_financial == 1 ~ 1,
      is.na(a19i_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # Why_Housing: Keep as is (already exists as a19i_housing)
    Why_Housing = case_when(
      a19i_housing == 1 ~ 1,
      is.na(a19i_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # Why_New_Environ: Keep as is (already exists as a19i_new_environment)
    Why_New_Environ = case_when(
      a19i_new_environment == 1 ~ 1,
      is.na(a19i_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # Why_Family_&_Friends: Family Support OR Meet Friends
    Why_Family_Friends = case_when(
      a19i_family_support == 1 | a19i_meet_friends == 1 ~ 1,
      is.na(a19i_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # Why_Conflict_&_Forced: Conflicts OR Forced
    Why_Conflict_Forced = case_when(
      a19i_conflicts == 1 | a19i_forced == 1 ~ 1,
      is.na(a19i_response) ~ NA_real_,
      TRUE ~ 0
    ),

    # Why_Other_&_Unclear: Other OR Not Clear
    Why_Other_Unclear = case_when(
      a19i_other == 1 | a19i_not_clear == 1 ~ 1,
      is.na(a19i_response) ~ NA_real_,
      TRUE ~ 0
    )
  )

cat("Created collapsed move reason categories:\n")
cat("  Why_Spo_Part_End     - Partner/Spouse + Relationship End\n")
cat("  Why_Own_Part_Career  - Own Job/School + Partner's Job/School\n")
cat("  Why_Financial        - Financial reasons\n")
cat("  Why_Housing          - Housing reasons\n")
cat("  Why_New_Environ      - New living environment\n")
cat("  Why_Family_Friends   - Family Support + Meet Friends\n")
cat("  Why_Conflict_Forced  - Conflicts + Forced to move\n")
cat("  Why_Other_Unclear    - Other + Not clear\n")
```

## Summary

```{r merge-summary}
cat("\n--- External Data Merge Summary ---\n")

cat("\nVariables added from a19f (Steps to Make Friends):\n")
cat("  a19f_response           - Open-ended response text\n")
cat("  a19f_religious          - Religious Community Involvement\n")
cat("  a19f_local_venue        - Local Venue Visitation\n")
cat("  a19f_volunteer          - Secular Volunteer Service\n")
cat("  a19f_online             - Online Social Networking\n")
cat("  a19f_workplace          - Workplace Socializing\n")
cat("  a19f_athletic           - Athletic Engagement\n")
cat("  a19f_cultural_hobby     - Cultural-Hobby Participation\n")
cat("  a19f_neighborhood       - Neighborhood Integration\n")
cat("  a19f_events_groups      - Broader Event and Group Participation\n")
cat("  a19f_through_connections - Friendship Through Connections\n")
cat("  a19f_non_engagement     - Non-Engagement Preference\n")
cat("  a19f_alternative        - Alternative Strategies\n")
cat("  a19f_not_clear          - Not Clear / Uncodeable Responses\n")
cat("  a19f_no_match           - No Match flag\n")

cat("\nCollapsed a19f categories (How_ prefix):\n")
cat("  GRANULAR:\n")
cat("    How_Workplace          - Workplace socializing\n")
cat("    How_Through_Connections - Through existing connections\n")
cat("    How_Community_Orgs     - Religious + Volunteer\n")
cat("    How_Activities         - Athletic + Cultural/Hobby + Events/Groups\n")
cat("    How_Local_Places       - Local Venue + Neighborhood\n")
cat("    How_Online             - Online social networking\n")
cat("    How_Non_Engagement     - Prefers not to seek new friends\n")
cat("    How_Other_Unclear      - Alternative + Not clear\n")
cat("  BROADER:\n")
cat("    How_Network_Ties       - Workplace + Through Connections + Religious\n")
cat("    How_Active_Outreach    - All proactive activity-based strategies\n")
cat("    How_Any_Engagement     - Any active strategy (binary)\n")

cat("\nVariables added from a19i (Reasons for Moving):\n")
cat("  a19i_response           - Open-ended response text\n")
cat("  a19i_partner_spouse     - Partner or spouse\n")
cat("  a19i_relationship_end   - Relationship end\n")
cat("  a19i_own_job_school     - Own job/school/career\n")
cat("  a19i_partner_job_school - Partner's job/school/career\n")
cat("  a19i_financial          - Financial\n")
cat("  a19i_housing            - Housing, concretely\n")
cat("  a19i_new_environment    - New living environment\n")
cat("  a19i_family_support     - Family Support\n")
cat("  a19i_meet_friends       - Meet and Make Friends\n")
cat("  a19i_conflicts          - Conflicts\n")
cat("  a19i_forced             - Forced to move\n")
cat("  a19i_other              - Other reason\n")
cat("  a19i_not_clear          - Not clear\n")
cat("  a19i_forced_financial   - Forced Financial\n")
cat("  a19i_not_forced_financial - Not Forced Financial\n")
cat("  a19i_financial_housing  - Financial Housing\n")
cat("  a19i_non_financial_housing - Non-Financial Housing\n")
cat("  a19i_financial_family_support - Financial Family Support\n")
cat("  a19i_non_financial_family_support - Non-Financial Family Support\n")

cat("\nExternal data merge complete.\n")
```
