---
title: "Move Reason Descriptives"
author: "Migration Networks UCNets Project"
date: "`r Sys.Date()`"
output:
  html_document:
    css: !expr here::here("code", "styles.css")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = TRUE, warning = TRUE)
knitr::opts_knit$set(root.dir = here::here())
options(width = 200)
```

```{r load-kable, include=FALSE}
library(knitr)
library(kableExtra)

# Helper function for styled tables
styled_table <- function(df, caption = NULL) {
  kable(df, row.names = FALSE, caption = caption) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "left"
    )
}
```

# Overview

This notebook analyzes reasons for moving using collapsed categories from the a19i open-ended responses about "why did you move?"

## Key Takeaways

- **Career drives young adult mobility**: Career-related reasons dominate among young movers but drop sharply for older adults, reflecting life stage differences in labor market attachment.
- **Financial pressures increase with age**: Older movers cite financial reasons at nearly twice the rate of young movers, suggesting economic constraints play a larger role later in life.
- **Family/friends as pull factors**: Movers who cite family/friends have significantly more kin ties, indicating these moves are drawn toward existing networks rather than away from current locations.
- **Education shapes move narratives**: College-educated movers emphasize career reasons, while those without degrees more often cite involuntary reasons like conflict or being forced to move.
- **Homeownership reflects housing-driven moves**: Homeowners who move disproportionately cite housing reasons, suggesting housing transitions (upsizing, downsizing) motivate their relocations.

## Category Mapping

The table below shows how original a19i response categories map to our collapsed categories:

| Collapsed Category | Original a19i Categories |
|-------------------|-------------------------|
| **Partner/Relationship** | Partner/Spouse + Relationship End |
| **Career (Own/Partner)** | Own Job/School + Partner's Job/School |
| **Financial** | Financial reasons |
| **Housing** | Housing reasons |
| **New Environment** | New living environment |
| **Family/Friends** | Family Support + Meet Friends |
| **Conflict/Forced** | Conflicts + Forced to move |
| **Other/Unclear** | Other + Not clear |

## Load Libraries and Helpers

```{r libraries}
library(here)
library(dplyr)
library(readxl)

# Source helper functions
source(here("code", "helpers", "descriptive_helpers.R"))
```

## Load Data

```{r load-data, include=FALSE}
df <- readRDS(here("data", "out", "R_DF.rds"))

# Create female indicator (gender: 1=Male, 2=Female)
df$female <- ifelse(df$gender == 2, 1, 0)

cat(sprintf("Loaded %d observations for %d unique individuals\n",
            nrow(df),
            n_distinct(df$prim_key)))
```

# Why Do People Move?

## Overall Percentages by Age Group

What percentage of movers reported each reason for moving?

```{r move-reasons-overall}
library(haven)

# Load a19i data
a19i_raw <- read_excel(here("data", "in", "Final_Merged_Codes_a19i_alternate.xlsx"))

# Rename columns by position to handle corrupted column names
new_names <- c("prim_key", "WAVE", "prim_key_wave", "a19i_response",
               "a19i_partner_spouse", "a19i_relationship_end", "a19i_own_job_school",
               "a19i_partner_job_school", "a19i_financial", "a19i_housing",
               "a19i_new_environment", "a19i_family_support", "a19i_meet_friends",
               "a19i_conflicts", "a19i_forced", "a19i_other", "a19i_not_clear",
               "a19i_forced_financial", "a19i_not_forced_financial",
               "a19i_financial_housing", "a19i_non_financial_housing",
               "a19i_financial_family_support", "a19i_non_financial_family_support")
names(a19i_raw) <- new_names

a19i_raw$prim_key_wave <- as.character(as.numeric(a19i_raw$prim_key_wave))

# Load main data to get agegroup
df_main <- read_dta(here("data", "R_DF.dta"))
df_main <- df_main %>%
  mutate(prim_key_wave = paste0(as.character(prim_key), WAVE)) %>%
  select(prim_key_wave, agegroup)

# Merge to get agegroup
a19i <- a19i_raw %>%
  left_join(df_main, by = "prim_key_wave")

n_young <- sum(a19i$agegroup == 0, na.rm = TRUE)
n_old <- sum(a19i$agegroup == 1, na.rm = TRUE)
n_total <- nrow(a19i)

cat(sprintf("Movers with a19i data: %d (Young: %d, Old: %d)\n", n_total, n_young, n_old))

# Define move reason categories
move_reasons <- c(
  "Partner/Relationship",
  "Career (Own/Partner)",
  "Financial",
  "Housing",
  "New Environment",
  "Family/Friends",
  "Conflict/Forced",
  "Other/Unclear"
)

# Helper function to calculate N for each move reason
calc_n_move <- function(data) {
  c(
    sum(data$a19i_partner_spouse == 1 | data$a19i_relationship_end == 1, na.rm = TRUE),
    sum(data$a19i_own_job_school == 1 | data$a19i_partner_job_school == 1, na.rm = TRUE),
    sum(data$a19i_financial == 1, na.rm = TRUE),
    sum(data$a19i_housing == 1, na.rm = TRUE),
    sum(data$a19i_new_environment == 1, na.rm = TRUE),
    sum(data$a19i_family_support == 1 | data$a19i_meet_friends == 1, na.rm = TRUE),
    sum(data$a19i_conflicts == 1 | data$a19i_forced == 1, na.rm = TRUE),
    sum(data$a19i_other == 1 | data$a19i_not_clear == 1, na.rm = TRUE)
  )
}

# Calculate for each age group
young_data <- a19i %>% filter(agegroup == 0)
old_data <- a19i %>% filter(agegroup == 1)

young_n <- calc_n_move(young_data)
old_n <- calc_n_move(old_data)
total_n <- calc_n_move(a19i)

# Build table
move_reasons_overall <- data.frame(
  Reason = move_reasons,
  Young_N = young_n,
  Young_Pct = sprintf("%.1f%%", 100 * young_n / n_young),
  Old_N = old_n,
  Old_Pct = sprintf("%.1f%%", 100 * old_n / n_old),
  Total_N = total_n,
  Total_Pct = sprintf("%.1f%%", 100 * total_n / n_total),
  stringsAsFactors = FALSE
)

# Add sample size row
sample_row <- data.frame(
  Reason = "N (movers)",
  Young_N = n_young,
  Young_Pct = "",
  Old_N = n_old,
  Old_Pct = "",
  Total_N = n_total,
  Total_Pct = "",
  stringsAsFactors = FALSE
)

move_reasons_overall <- bind_rows(move_reasons_overall, sample_row)

styled_table(move_reasons_overall, "Move Reasons by Age Group")

export_descriptive_table(move_reasons_overall, "descriptives_move_reasons_overall.csv")
```

```{r move-reasons-overall-plot, fig.width=10, fig.height=6}
library(ggplot2)
library(tidyr)

# Prepare data for plotting (exclude N row and Other/Unclear)
plot_reasons_data <- move_reasons_overall %>%
  filter(Reason != "N (movers)", Reason != "Other/Unclear") %>%
  select(Reason, Young_N, Old_N) %>%
  mutate(
    Young_Pct = 100 * Young_N / n_young,
    Old_Pct = 100 * Old_N / n_old
  ) %>%
  select(Reason, Young_Pct, Old_Pct) %>%
  pivot_longer(cols = c(Young_Pct, Old_Pct), names_to = "Age_Group", values_to = "Percentage") %>%
  mutate(
    Age_Group = ifelse(Age_Group == "Young_Pct", "Young (21-30)", "Old (50-70)"),
    # Order reasons by total usage
    Reason = factor(Reason, levels = c("Career (Own/Partner)", "Financial", "Housing",
                                        "Partner/Relationship", "Conflict/Forced",
                                        "New Environment", "Family/Friends"))
  )

ggplot(plot_reasons_data, aes(x = Reason, y = Percentage, fill = Age_Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  scale_fill_manual(values = c("Young (21-30)" = "#2166AC", "Old (50-70)" = "#B2182B")) +
  labs(
    title = "Reasons for Moving by Age Group",
    subtitle = "Percentage of movers reporting each reason",
    x = "",
    y = "Percentage",
    fill = "Age Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 25, hjust = 1, size = 10),
    plot.title = element_text(face = "bold", size = 14),
    panel.grid.major.x = element_blank()
  ) +
  coord_cartesian(ylim = c(0, 40))
```

**Key patterns:** Career reasons are the most common motivation for moving overall (30.2%), but this is driven almost entirely by young adults (34.6% vs. 12.0% old). In contrast, financial reasons are cited nearly twice as often by older movers (27.7% vs. 15.2%), and family/friends as pull factors are also more common among older adults (18.1% vs. 8.4%). Partner/relationship reasons are somewhat more common among young movers (15.7% vs. 9.6%). Housing and conflict/forced reasons show more similar rates across age groups.

## By Demographic Group

Percentages by demographic subgroups (columns show % within each group).

```{r move-reasons-by-demo}
# Get additional demographic variables from main data
df_demo <- read_dta(here("data", "R_DF.dta"))
df_demo <- df_demo %>%
  mutate(prim_key_wave = paste0(as.character(prim_key), WAVE)) %>%
  select(prim_key_wave, agegroup, gender, Marr_Cohab, Eth_black, Eth_asian, Eth_latino)

# Create female indicator
df_demo$female <- ifelse(df_demo$gender == 2, 1, 0)

# Merge with a19i
a19i_demo <- a19i_raw %>%
  left_join(df_demo, by = "prim_key_wave")

# Define demographic groups
demo_groups <- list(
  "Young" = list(var = "agegroup", val = 0),
  "Old" = list(var = "agegroup", val = 1),
  "Male" = list(var = "female", val = 0),
  "Female" = list(var = "female", val = 1),
  "Not Married" = list(var = "Marr_Cohab", val = 0),
  "Married" = list(var = "Marr_Cohab", val = 1),
  "Black" = list(var = "Eth_black", val = 1),
  "Asian" = list(var = "Eth_asian", val = 1),
  "Latino" = list(var = "Eth_latino", val = 1)
)

# Build results table
move_reasons_demo <- data.frame(Reason = move_reasons, stringsAsFactors = FALSE)

for (grp_name in names(demo_groups)) {
  grp <- demo_groups[[grp_name]]
  grp_data <- a19i_demo %>% filter(.data[[grp$var]] == grp$val)
  n_grp <- nrow(grp_data)

  grp_n <- calc_n_move(grp_data)
  move_reasons_demo[[grp_name]] <- sprintf("%.1f%%", 100 * grp_n / n_grp)
}

# Add Total column
move_reasons_demo$Total <- sprintf("%.1f%%", 100 * total_n / n_total)

# Add N row
n_row <- data.frame(Reason = "N", stringsAsFactors = FALSE)
for (grp_name in names(demo_groups)) {
  grp <- demo_groups[[grp_name]]
  n_row[[grp_name]] <- as.character(sum(a19i_demo[[grp$var]] == grp$val, na.rm = TRUE))
}
n_row$Total <- as.character(n_total)

move_reasons_demo <- bind_rows(move_reasons_demo, n_row)

styled_table(move_reasons_demo, "Move Reasons by Demographic Group")

export_descriptive_table(move_reasons_demo, "descriptives_move_reasons_demo.csv")
```

```{r move-reasons-demo-plot, fig.width=10, fig.height=8}
# Prepare data for plotting (Gender and Race/Ethnicity groups only)
plot_reasons_demo_data <- move_reasons_demo %>%
  filter(Reason != "N", Reason != "Other/Unclear") %>%
  select(Reason, Male, Female, Black, Asian, Latino) %>%
  mutate(across(-Reason, ~as.numeric(gsub("%", "", .x)))) %>%
  pivot_longer(cols = -Reason, names_to = "Group", values_to = "Percentage") %>%
  mutate(
    Category = case_when(
      Group %in% c("Male", "Female") ~ "Gender",
      Group %in% c("Black", "Asian", "Latino") ~ "Race/Ethnicity"
    ),
    # Order reasons by total usage
    Reason = factor(Reason, levels = c("Career (Own/Partner)", "Financial", "Housing",
                                        "Partner/Relationship", "Conflict/Forced",
                                        "New Environment", "Family/Friends")),
    # Set factor levels for groups
    Group = factor(Group, levels = c("Male", "Female", "Black", "Asian", "Latino"))
  )

ggplot(plot_reasons_demo_data, aes(x = Reason, y = Percentage, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  facet_wrap(~Category, ncol = 1, scales = "free_x") +
  scale_fill_manual(values = c(
    "Male" = "#2166AC",
    "Female" = "#B2182B",
    "Black" = "#1B9E77",
    "Asian" = "#D95F02",
    "Latino" = "#7570B3"
  )) +
  labs(
    title = "Reasons for Moving by Demographics",
    subtitle = "Percentage of movers reporting each reason",
    x = "",
    y = "Percentage",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 25, hjust = 1, size = 10),
    plot.title = element_text(face = "bold", size = 14),
    panel.grid.major.x = element_blank(),
    strip.text = element_text(face = "bold", size = 11)
  ) +
  coord_cartesian(ylim = c(0, 45))
```

**Key patterns:** Marital status strongly shapes move reasons. Married movers cite partner/relationship reasons at over three times the rate of unmarried movers (23.3% vs. 6.9%) and housing reasons nearly twice as often (21.8% vs. 11.6%). Unmarried movers emphasize career (33.5%) and conflict (16.3%) more. Gender differences are notable for family/friends (13.1% women vs. 4.3% men) and conflict (16.1% women vs. 9.2% men).

## By Socioeconomic Status

Percentages by education, homeownership, and employment status.

```{r move-reasons-by-ses}
# Get SES variables from main data
df_ses <- read_dta(here("data", "R_DF.dta"))
df_ses <- df_ses %>%
  mutate(prim_key_wave = paste0(as.character(prim_key), WAVE)) %>%
  select(prim_key_wave, LT_BA, Owns_home, Full_Time)

# Create BA+ indicator (inverse of LT_BA)
df_ses$BA_plus <- ifelse(df_ses$LT_BA == 1, 0, 1)

# Merge with a19i
a19i_ses <- a19i_raw %>%
  left_join(df_ses, by = "prim_key_wave")

# Define SES groups (showing both categories for each variable)
ses_groups <- list(
  "BA+" = list(var = "BA_plus", val = 1),
  "No BA" = list(var = "BA_plus", val = 0),
  "Homeowner" = list(var = "Owns_home", val = 1),
  "Renter" = list(var = "Owns_home", val = 0),
  "Full-time" = list(var = "Full_Time", val = 1),
  "Not FT" = list(var = "Full_Time", val = 0)
)

# Build results table
move_reasons_ses <- data.frame(Reason = move_reasons, stringsAsFactors = FALSE)

for (grp_name in names(ses_groups)) {
  grp <- ses_groups[[grp_name]]
  grp_data <- a19i_ses %>% filter(.data[[grp$var]] == grp$val)
  n_grp <- nrow(grp_data)

  grp_n <- calc_n_move(grp_data)
  move_reasons_ses[[grp_name]] <- sprintf("%.1f%%", 100 * grp_n / n_grp)
}

# Add Total column
move_reasons_ses$Total <- sprintf("%.1f%%", 100 * total_n / n_total)

# Add N row
n_row_ses <- data.frame(Reason = "N", stringsAsFactors = FALSE)
for (grp_name in names(ses_groups)) {
  grp <- ses_groups[[grp_name]]
  n_row_ses[[grp_name]] <- as.character(sum(a19i_ses[[grp$var]] == grp$val, na.rm = TRUE))
}
n_row_ses$Total <- as.character(n_total)

move_reasons_ses <- bind_rows(move_reasons_ses, n_row_ses)

styled_table(move_reasons_ses, "Move Reasons by Socioeconomic Status")

export_descriptive_table(move_reasons_ses, "descriptives_move_reasons_ses.csv")
```

**Key patterns:** Education creates stark differences in move reasons. BA+ holders cite career reasons at twice the rate of those without degrees (33.2% vs. 15.5%), while those without a BA cite conflict/forced reasons at more than double the rate (26.8% vs. 11.4%). Homeownership also shapes move narratives: homeowners who move cite housing reasons at nearly three times the rate of renters (31.0% vs. 12.1%) and family/friends more often (18.0% vs. 8.0%), while renters cite conflict more (15.6% vs. 8.0%).

```{r move-reasons-ses-plot, fig.width=12, fig.height=10}
library(ggplot2)
library(tidyr)

# Prepare data for plotting (exclude N row and Total column)
plot_data_ses <- move_reasons_ses %>%
  filter(Reason != "N") %>%
  select(Reason, `BA+`, `No BA`, Homeowner, Renter, `Full-time`, `Not FT`) %>%
  pivot_longer(cols = -Reason, names_to = "Group", values_to = "Percent") %>%
  mutate(
    Percent = as.numeric(gsub("%", "", Percent)),
    SES_Dimension = case_when(
      Group %in% c("BA+", "No BA") ~ "Education",
      Group %in% c("Homeowner", "Renter") ~ "Homeownership",
      Group %in% c("Full-time", "Not FT") ~ "Employment"
    ),
    SES_Dimension = factor(SES_Dimension, levels = c("Education", "Homeownership", "Employment")),
    Group = factor(Group, levels = c("BA+", "No BA", "Homeowner", "Renter", "Full-time", "Not FT")),
    Reason = factor(Reason, levels = c("Partner/Relationship", "Career (Own/Partner)",
                                        "Financial", "Housing", "New Environment",
                                        "Family/Friends", "Conflict/Forced", "Other/Unclear"))
  )

# Create faceted bar plot
ggplot(plot_data_ses, aes(x = Reason, y = Percent, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  facet_wrap(~SES_Dimension, ncol = 1) +
  scale_fill_manual(values = c("BA+" = "#2166AC", "No BA" = "#B2182B",
                                "Homeowner" = "#1B7837", "Renter" = "#762A83",
                                "Full-time" = "#E66101", "Not FT" = "#E91E63"),
                    name = "Group") +
  labs(
    title = "Move Reasons by Socioeconomic Status",
    subtitle = "Percentage of movers citing each reason, by education, homeownership, and employment",
    x = "",
    y = "Percent"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11, face = "italic"),
    legend.position = "right",
    legend.title = element_text(size = 10, face = "bold"),
    strip.text = element_text(size = 12, face = "bold"),
    panel.grid.major.x = element_blank()
  ) +
  scale_y_continuous(labels = function(x) paste0(x, "%"), limits = c(0, 40))
```

## By Network Size

Average network size for those who cite vs. don't cite each move reason.

This table compares mean network sizes between those who cite each reason and those who don't. "NonKin_Use" shows the average number of non-kin ties for movers who cite that reason, while "NonKin_NoUse" shows the average for those who don't. The "sig" columns indicate whether the difference is statistically significant based on t-tests (*** p<.001, ** p<.01, * p<.05). The same pattern applies to kin ties.

```{r move-reasons-by-network-avg}
# Get network variables from main data
df_net <- read_dta(here("data", "R_DF.dta"))
df_net <- df_net %>%
  mutate(
    prim_key_wave = paste0(as.character(prim_key), WAVE),
    TotalKin = ImmedKin + ExtKin
  ) %>%
  select(prim_key_wave, NonKin, TotalKin)

# Merge with a19i
a19i_net <- a19i_raw %>%
  left_join(df_net, by = "prim_key_wave")

cat(sprintf("Overall means: NonKin = %.2f, Kin = %.2f\n\n",
            mean(a19i_net$NonKin, na.rm = TRUE),
            mean(a19i_net$TotalKin, na.rm = TRUE)))

# Move reason variable mappings
reason_vars <- list(
  "Partner/Relationship" = c("a19i_partner_spouse", "a19i_relationship_end"),
  "Career (Own/Partner)" = c("a19i_own_job_school", "a19i_partner_job_school"),
  "Financial" = "a19i_financial",
  "Housing" = "a19i_housing",
  "New Environment" = "a19i_new_environment",
  "Family/Friends" = c("a19i_family_support", "a19i_meet_friends"),
  "Conflict/Forced" = c("a19i_conflicts", "a19i_forced"),
  "Other/Unclear" = c("a19i_other", "a19i_not_clear")
)

# Helper function to check if reason is cited
cites_reason <- function(data, vars) {
  if (length(vars) == 1) {
    return(data[[vars]] == 1)
  } else {
    result <- rep(FALSE, nrow(data))
    for (v in vars) {
      result <- result | (data[[v]] == 1)
    }
    return(result)
  }
}

# Function to format p-value with stars
format_p <- function(p) {
  if (is.na(p)) return("")
  if (p < 0.001) return("***")
  if (p < 0.01) return("**")
  if (p < 0.05) return("*")
  return("")
}

# Calculate mean network size for those who cite vs don't cite each reason with t-tests
move_reasons_net_avg <- data.frame(Reason = character(), stringsAsFactors = FALSE)

for (reason_name in names(reason_vars)) {
  vars <- reason_vars[[reason_name]]

  cite_idx <- cites_reason(a19i_net, vars)
  citers <- a19i_net[cite_idx, ]
  non_citers <- a19i_net[!cite_idx, ]

  # T-tests for NonKin and Kin
  ttest_nonkin <- t.test(citers$NonKin, non_citers$NonKin)
  ttest_kin <- t.test(citers$TotalKin, non_citers$TotalKin)

  move_reasons_net_avg <- rbind(move_reasons_net_avg, data.frame(
    Reason = reason_name,
    N_Cite = nrow(citers),
    NonKin_Cite = sprintf("%.2f", mean(citers$NonKin, na.rm = TRUE)),
    NonKin_NoCite = sprintf("%.2f", mean(non_citers$NonKin, na.rm = TRUE)),
    NonKin_sig = format_p(ttest_nonkin$p.value),
    Kin_Cite = sprintf("%.2f", mean(citers$TotalKin, na.rm = TRUE)),
    Kin_NoCite = sprintf("%.2f", mean(non_citers$TotalKin, na.rm = TRUE)),
    Kin_sig = format_p(ttest_kin$p.value),
    stringsAsFactors = FALSE
  ))
}

styled_table(move_reasons_net_avg, "Average Network Size by Move Reason (t-test: *** p<.001, ** p<.01, * p<.05)")

export_descriptive_table(move_reasons_net_avg, "descriptives_move_reasons_network_avg.csv")
```

**Key patterns:** Movers who cite family/friends as a reason have significantly more kin ties than those who don't (3.36 vs. 2.64*), consistent with these being pull moves toward existing family networks. Movers citing housing reasons have larger non-kin networks (8.06 vs. 7.05), while career movers also show larger non-kin networks (7.77 vs. 6.98). Those citing new environment or other/unclear reasons have smaller non-kin networks, though not statistically significant.

## By Network Size (Terciles)

Percentages citing each reason within network size terciles (Low/Medium/High).

```{r move-reasons-by-network-terciles}
# Create terciles for NonKin
nonkin_terciles <- quantile(a19i_net$NonKin, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE)
a19i_net <- a19i_net %>%
  mutate(
    NonKin_tercile = case_when(
      NonKin <= nonkin_terciles[2] ~ "Low",
      NonKin <= nonkin_terciles[3] ~ "Med",
      TRUE ~ "High"
    ),
    NonKin_tercile = factor(NonKin_tercile, levels = c("Low", "Med", "High"))
  )

# Create terciles for Kin
kin_terciles <- quantile(a19i_net$TotalKin, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE)
a19i_net <- a19i_net %>%
  mutate(
    Kin_tercile = case_when(
      TotalKin <= kin_terciles[2] ~ "Low",
      TotalKin <= kin_terciles[3] ~ "Med",
      TRUE ~ "High"
    ),
    Kin_tercile = factor(Kin_tercile, levels = c("Low", "Med", "High"))
  )

cat("NonKin tercile cutpoints:", nonkin_terciles, "\n")
cat("Kin tercile cutpoints:", kin_terciles, "\n\n")

# Define network groups
net_groups <- list(
  "NonKin Low" = list(var = "NonKin_tercile", val = "Low"),
  "NonKin Med" = list(var = "NonKin_tercile", val = "Med"),
  "NonKin High" = list(var = "NonKin_tercile", val = "High"),
  "Kin Low" = list(var = "Kin_tercile", val = "Low"),
  "Kin Med" = list(var = "Kin_tercile", val = "Med"),
  "Kin High" = list(var = "Kin_tercile", val = "High")
)

# Build results table
move_reasons_net <- data.frame(Reason = move_reasons, stringsAsFactors = FALSE)

for (grp_name in names(net_groups)) {
  grp <- net_groups[[grp_name]]
  grp_data <- a19i_net %>% filter(.data[[grp$var]] == grp$val)
  n_grp <- nrow(grp_data)

  grp_n <- calc_n_move(grp_data)
  move_reasons_net[[grp_name]] <- sprintf("%.1f%%", 100 * grp_n / n_grp)
}

# Add Total column
move_reasons_net$Total <- sprintf("%.1f%%", 100 * total_n / n_total)

# Add N row
n_row_net <- data.frame(Reason = "N", stringsAsFactors = FALSE)
for (grp_name in names(net_groups)) {
  grp <- net_groups[[grp_name]]
  n_row_net[[grp_name]] <- as.character(sum(a19i_net[[grp$var]] == grp$val, na.rm = TRUE))
}
n_row_net$Total <- as.character(n_total)

move_reasons_net <- bind_rows(move_reasons_net, n_row_net)

styled_table(move_reasons_net, "Move Reasons by Network Size (Terciles)")

export_descriptive_table(move_reasons_net, "descriptives_move_reasons_network_terciles.csv")
```

**Key patterns:** Family/friends as a move reason increases steadily across kin network terciles (7.6% low to 14.8% high), reinforcing that these are pull moves toward family. Housing reasons increase across non-kin terciles (11.1% to 21.3%), as do career reasons (27.2% to 34.6%), suggesting movers with larger friend networks may be more embedded in professional contexts. Those with smaller non-kin networks more often cite family/friends (12.3% vs. 7.9% for high tercile).

```{r move-reasons-terciles-plot, fig.width=12, fig.height=8}
library(ggplot2)
library(tidyr)

# Prepare data for plotting (exclude N row)
plot_data_net <- move_reasons_net %>%
  filter(Reason != "N") %>%
  select(Reason, `NonKin Low`, `NonKin Med`, `NonKin High`,
         `Kin Low`, `Kin Med`, `Kin High`) %>%
  pivot_longer(cols = -Reason, names_to = "Group", values_to = "Percent") %>%
  mutate(
    Percent = as.numeric(gsub("%", "", Percent)),
    Network_Type = ifelse(grepl("NonKin", Group), "Non-Kin Network", "Kin Network"),
    Tercile = case_when(
      grepl("Low", Group) ~ "Low",
      grepl("Med", Group) ~ "Medium",
      grepl("High", Group) ~ "High"
    ),
    Tercile = factor(Tercile, levels = c("Low", "Medium", "High")),
    Reason = factor(Reason, levels = c("Partner/Relationship", "Career (Own/Partner)",
                                        "Financial", "Housing", "New Environment",
                                        "Family/Friends", "Conflict/Forced", "Other/Unclear"))
  )

# Create faceted bar plot
ggplot(plot_data_net, aes(x = Reason, y = Percent, fill = Tercile)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  facet_wrap(~Network_Type, ncol = 1) +
  scale_fill_manual(values = c("Low" = "#2166AC", "Medium" = "#F4A582", "High" = "#B2182B"),
                    name = "Network Size\nTercile") +
  labs(
    title = "Move Reasons by Network Size Terciles",
    subtitle = "Percentage of movers citing each reason, by non-kin and kin network size",
    x = "",
    y = "Percent"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11, face = "italic"),
    legend.position = "right",
    legend.title = element_text(size = 10, face = "bold"),
    strip.text = element_text(size = 12, face = "bold"),
    panel.grid.major.x = element_blank()
  ) +
  scale_y_continuous(labels = function(x) paste0(x, "%"), limits = c(0, 40))
```
