---
title: "Fixed Effects Panel Models (Young Adults)"
author: "Migration Networks UCNets Project"
date: "`r Sys.Date()`"
output:
  html_document:
    css: !expr here::here("code", "styles.css")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

# Overview

This notebook runs **fixed effects panel models** with robust standard errors for the Migration Networks analysis. Fixed effects models estimate within-person effects by differencing out all time-invariant characteristics.

**Models focus on young adults (ages 21-30) only.**

## Key Takeaways

> **1. Focus on young adults only.** Older adults were dropped due to severe power limitations (EPV = 1.1 for out-of-Bay moves with only 30 events).
>
> **2. Fixed effects models estimate within-person effects.** These are the causal estimates of interest - how changes in predictors for the *same person* relate to mobility outcomes.
>
> **3. Only time-varying predictors are included.** Demographics (race, sex) and between-person means are absorbed by the fixed effect and cannot be estimated.
>
> **4. Power is improved vs. mixed models.** With 15 predictors instead of 27, EPV improves substantially.
>
> **5. Robust standard errors (HC1) account for heteroskedasticity.**

## Statistical Power Summary

| Outcome | Events | Predictors | EPV | Status |
|---------|--------|------------|-----|--------|
| Out-of-Bay | 114 | 15 | 7.6 | Marginal |
| Within-Bay | 235 | 15 | 15.7 | OK |
| Plan to Move | 379 | 15 | 25.3 | OK |

**Note:** EPV ≥10 is the guideline for stable estimates. Out-of-Bay is marginal but much improved from the mixed model (EPV = 4.2).

## Load Libraries and Helpers

```{r libraries}
library(here)
library(dplyr)

# Source helper functions
source(here("code", "helpers", "fe_model_helpers.R"))
```

## Load Data

```{r load-data}
df <- readRDS(here("data", "out", "R_DF.rds"))

cat(sprintf("Loaded %d observations for %d unique individuals\n",
            nrow(df),
            n_distinct(df$prim_key)))

# Filter to young adults only
df_young <- df %>% filter(agegroup == 0)

cat(sprintf("\nYoung adults (agegroup=0): %d observations for %d individuals\n",
            nrow(df_young),
            n_distinct(df_young$prim_key)))

# Show outcome event rates
cat("\nOutcome event rates (young adults):\n")
cat(sprintf("  Out-of-Bay moves: %d (%.1f%%)\n",
            sum(df_young$Moved_out_of_Bay_w23, na.rm = TRUE),
            100 * mean(df_young$Moved_out_of_Bay_w23, na.rm = TRUE)))
cat(sprintf("  Within-Bay moves: %d (%.1f%%)\n",
            sum(df_young$Moved_in_Bay_w23, na.rm = TRUE),
            100 * mean(df_young$Moved_in_Bay_w23, na.rm = TRUE)))
cat(sprintf("  Plan to Move:     %d (%.1f%%)\n",
            sum(df_young$Plan_to_Move, na.rm = TRUE),
            100 * mean(df_young$Plan_to_Move, na.rm = TRUE)))
```

# Predictor Configuration

Fixed effects models use only **time-varying predictors**. Demographics and between-person means are absorbed by the individual fixed effect.

```{r predictors}
# Time-varying predictors (lagged)
# These are the ORIGINAL lagged variables, not the within-person deviations
# The FE model will automatically demean these
fe_predictors <- c(
  # Prior mobility
  "lag_Moved_Town",
  # Housing & employment
  "lag_Owns_Home",
  "lag_Full_Time",
  "lag_BA_or_greater",
  # Life events
  "lag_New_Job",
  "lag_New_School",
  "lag_Finan_Prob",
  "lag_LEX_Birth",
  # Partnership
  "lag_Marr_Cohab",
  # Network variables (CORE predictors)
  "lag_Kin_within_60_N",
  "lag_Nonkin_within_60_N",
  "lag_Nonkin_in_HH_N",
  "lag_Kin_in_HH_N",
  "lag_Kin_GT_Hour",
  "lag_Nonkin_GT_Hour"
)

# Plan predictor (added for "with plan" models)
plan_predictor <- "lag_Plan_Move"

cat("Predictor count:\n")
cat("  Time-varying predictors:", length(fe_predictors), "\n")
cat("  Plan predictor:", plan_predictor, "\n")
cat("\nNote: Demographics (race, sex, age) are absorbed by the fixed effect.\n")
cat("Between-person means are not needed - FE handles within-person estimation.\n")
```

# Model Configuration

Define all models in a single configuration list. Each model specifies:

- `name`: Model identifier
- `outcome`: Outcome variable
- `age_group`: 0 = young (all models use young adults only)
- `include_plan`: Whether to include plan predictor
- `exclude_out_of_bay`: Whether to exclude out-of-bay movers (for within-bay models)

```{r model-specs}
model_specs <- list(
  # Out-of-Bay models (EPV = 7.6, marginal)
  list(name = "out_bay_no_plan",
       outcome = "Moved_out_of_Bay_w23",
       age_group = 0,
       include_plan = FALSE,
       exclude_out_of_bay = FALSE),

  list(name = "out_bay_with_plan",
       outcome = "Moved_out_of_Bay_w23",
       age_group = 0,
       include_plan = TRUE,
       exclude_out_of_bay = FALSE),

  # Plan to Move model (EPV = 25.3, OK)
  list(name = "plan_to_move",
       outcome = "Plan_to_Move",
       age_group = 0,
       include_plan = FALSE,
       exclude_out_of_bay = FALSE),

  # Within-Bay models (EPV = 15.7, OK)
  list(name = "within_bay_no_plan",
       outcome = "Moved_in_Bay_w23",
       age_group = 0,
       include_plan = FALSE,
       exclude_out_of_bay = TRUE),

  list(name = "within_bay_with_plan",
       outcome = "Moved_in_Bay_w23",
       age_group = 0,
       include_plan = TRUE,
       exclude_out_of_bay = TRUE)
)

cat("Total models to run:", length(model_specs), "\n")
cat("\nNote: All models use young adults (agegroup = 0) only.\n")
cat("Older adults excluded due to insufficient events (EPV < 2).\n")
```

# Run Models

```{r run-models}
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("RUNNING FIXED EFFECTS MODELS (YOUNG ADULTS ONLY)\n")
cat(paste(rep("=", 70), collapse = ""), "\n")

all_results <- run_all_fe_models(model_specs, df, fe_predictors, plan_predictor)
```

# Export Results

```{r export-results}
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("EXPORTING RESULTS\n")
cat(paste(rep("=", 70), collapse = ""), "\n")

# Out of Bay results (young adults only)
export_fe_results(
  list(all_results$out_bay_no_plan,
       all_results$out_bay_with_plan),
  c("No_Plan", "With_Plan"),
  "out_of_bay_moves_young_fe.csv"
)

# Plan to Move results (young adults only)
export_fe_results(
  list(all_results$plan_to_move),
  c("Young"),
  "plan_to_move_young_fe.csv"
)

# Within Bay results (young adults only)
export_fe_results(
  list(all_results$within_bay_no_plan,
       all_results$within_bay_with_plan),
  c("No_Plan", "With_Plan"),
  "within_bay_moves_young_fe.csv"
)
```

# Model Summaries

```{r summaries}
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("MODEL SUMMARIES\n")
cat(paste(rep("=", 70), collapse = ""), "\n")

print_all_fe_summaries(all_results)
```

# Save Model Objects

```{r save-models}
models <- extract_fe_model_objects(all_results)
saveRDS(models, here("out", "fe_models_young.rds"))
cat("\nModel objects saved to out/fe_models_young.rds\n")
```

# Summary

```{r final-summary}
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("FIXED EFFECTS MODELS COMPLETE (YOUNG ADULTS)\n")
cat(paste(rep("=", 70), collapse = ""), "\n")

cat("\nModels run:", length(all_results), "\n")
cat("Sample: Young adults (ages 21-30) only\n")
cat("Predictors:", length(fe_predictors), "time-varying variables\n")
cat("\nOutput files:\n")
cat("  - out/out_of_bay_moves_young_fe.csv\n")
cat("  - out/plan_to_move_young_fe.csv\n")
cat("  - out/within_bay_moves_young_fe.csv\n")
cat("  - out/fe_models_young.rds\n")

cat("\n", paste(rep("-", 70), collapse = ""), "\n")
cat("STATISTICAL POWER NOTES\n")
cat(paste(rep("-", 70), collapse = ""), "\n")
cat("
With 15 predictors (down from 27 in mixed models):

- Out-of-Bay models: EPV = 114/15 = 7.6 (marginal, improved from 4.2)
- Within-Bay models: EPV = 235/15 = 15.7 (adequate)
- Plan to Move model: EPV = 379/15 = 25.3 (adequate)

Fixed effects advantages:
- Automatic within-person estimation (no Mundlak decomposition needed)
- Fewer parameters = better power
- Clean causal interpretation of within-person effects

Trade-off:
- Cannot estimate effects of time-invariant variables (demographics)
- For between-person effects, run separate random effects models
")
```
