---
title: "Validation: Fixed Effects vs Mixed Model Equivalence"
author: "Migration Networks UCNets Project"
date: "`r Sys.Date()`"
output:
  html_document:
    css: !expr here::here("code", "styles.css")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = here::here())
```

# Overview

This notebook validates that our mixed effects models with within/between decomposition produce within-person coefficients equivalent to fixed effects estimates. This is a **sanity check** for internal use, not for the paper.

## Key Takeaways

> **1. The Mundlak approach works.** When within/between decomposition is done correctly, the within-person coefficient from a mixed model equals the fixed effects coefficient.
>
> **2. Decomposition must match the analysis sample.** Pre-computed within/between variables (calculated on the full dataset) won't produce exact FE equivalence when the model uses a filtered subset.
>
> **3. Our Script 10 models use pre-computed decomposition.** This means within coefficients are close to but not identical to FE. This is acceptable for our purposes but worth documenting.
>
> **4. For exact equivalence, recompute on the filtered sample.** This validation shows how to do it correctly.

## The Mundlak/Chamberlain Approach

In panel data, we often want to separate:

- **Within-person effects**: How does a change in X for the *same person* relate to a change in Y?
- **Between-person effects**: How do people with different average X differ in average Y?

**Fixed effects** (FE) models estimate within-person effects by "demeaning" the data:

$$Y_{it} - \bar{Y}_i = \beta(X_{it} - \bar{X}_i) + \epsilon_{it}$$

**The Mundlak approach** includes both within-person deviations AND between-person means in a mixed model:

$$Y_{it} = \beta_{within}(X_{it} - \bar{X}_i) + \beta_{between}\bar{X}_i + \alpha_i + \epsilon_{it}$$

**Key result:** $\beta_{within}$ from the Mundlak model equals $\beta$ from the FE model.

## Load Libraries

```{r libraries}
library(here)
library(dplyr)
library(lme4)
library(plm)
library(knitr)
library(kableExtra)

styled_table <- function(df, caption = NULL) {
  kable(df, row.names = FALSE, caption = caption, digits = 6) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "left"
    )
}
```

## Load Data

```{r load-data}
df <- readRDS(here("data", "out", "R_DF.rds"))
cat(sprintf("Full dataset: %d observations, %d individuals\n", nrow(df), n_distinct(df$prim_key)))
```

# Test 1: Using Pre-Computed Decomposition (Current Approach)

The dataset has pre-computed `within_*` and `between_*` variables calculated on the **full panel**. Let's see what happens when we use these on a filtered subset.

```{r test1-precomputed}
# Filter to young adults with complete data
test_data <- df %>%
  filter(agegroup == 0) %>%
  filter(!is.na(Moved_out_of_Bay_w23),
         !is.na(within_lag_Kin_within_60_N),
         !is.na(between_lag_Kin_within_60_N))

cat(sprintf("Filtered sample: %d observations\n\n", nrow(test_data)))

# Fixed Effects model
pdata <- pdata.frame(test_data, index = c("prim_key", "WAVE"))
fe_model <- plm(Moved_out_of_Bay_w23 ~ lag_Kin_within_60_N,
                data = pdata, model = "within")
fe_coef <- coef(fe_model)["lag_Kin_within_60_N"]

# Mixed model with pre-computed decomposition
mixed_precomputed <- lmer(Moved_out_of_Bay_w23 ~ within_lag_Kin_within_60_N +
                          between_lag_Kin_within_60_N + (1 | prim_key),
                          data = test_data, REML = TRUE)
within_coef_pre <- fixef(mixed_precomputed)["within_lag_Kin_within_60_N"]

# Comparison
test1_results <- data.frame(
  Method = c("Fixed Effects (plm)", "Mixed with Pre-computed Decomposition"),
  Coefficient = c(fe_coef, within_coef_pre),
  stringsAsFactors = FALSE
)
test1_results$Difference_from_FE <- test1_results$Coefficient - fe_coef

styled_table(test1_results, "Test 1: Pre-computed Decomposition (Expect Mismatch)")
```

**Result:** The coefficients do NOT match because the within/between variables were computed on the full dataset, not the filtered subset.

# Test 2: Recomputing Decomposition on Analysis Sample

Now let's compute the within/between decomposition on the **same sample** used in the model.

```{r test2-recomputed}
# Recompute within/between on the filtered sample
test_data_recomputed <- test_data %>%
  group_by(prim_key) %>%
  mutate(
    person_mean = mean(lag_Kin_within_60_N, na.rm = TRUE),
    within_recomputed = lag_Kin_within_60_N - person_mean
  ) %>%
  ungroup()

# Mixed model with recomputed decomposition
mixed_recomputed <- lmer(Moved_out_of_Bay_w23 ~ within_recomputed + person_mean + (1 | prim_key),
                         data = test_data_recomputed, REML = TRUE)
within_coef_re <- fixef(mixed_recomputed)["within_recomputed"]

# Comparison
test2_results <- data.frame(
  Method = c("Fixed Effects (plm)",
             "Mixed with Pre-computed Decomposition",
             "Mixed with Recomputed Decomposition"),
  Coefficient = c(fe_coef, within_coef_pre, within_coef_re),
  stringsAsFactors = FALSE
)
test2_results$Difference_from_FE <- test2_results$Coefficient - fe_coef

styled_table(test2_results, "Test 2: Recomputed Decomposition (Expect Match)")
```

**Result:** The recomputed decomposition matches FE exactly.

# Test 3: Balanced Panel Validation

For maximum rigor, let's test on a balanced panel (individuals with data in all waves).

```{r test3-balanced}
# Create balanced panel
balanced <- df %>%
  filter(agegroup == 0, WAVE %in% c(2, 3)) %>%
  filter(!is.na(Moved_out_of_Bay_w23), !is.na(lag_Kin_within_60_N)) %>%
  group_by(prim_key) %>%
  filter(n() == 2) %>%
  ungroup()

cat(sprintf("Balanced panel: %d observations, %d individuals\n\n",
            nrow(balanced), n_distinct(balanced$prim_key)))

# Compute decomposition on balanced panel
balanced <- balanced %>%
  group_by(prim_key) %>%
  mutate(
    person_mean = mean(lag_Kin_within_60_N, na.rm = TRUE),
    within_manual = lag_Kin_within_60_N - person_mean
  ) %>%
  ungroup()

# Method 1: Fixed Effects (plm)
pdata_bal <- pdata.frame(balanced, index = c("prim_key", "WAVE"))
fe_balanced <- plm(Moved_out_of_Bay_w23 ~ lag_Kin_within_60_N,
                   data = pdata_bal, model = "within")

# Method 2: Mixed Effects (REML)
mixed_reml <- lmer(Moved_out_of_Bay_w23 ~ within_manual + person_mean + (1 | prim_key),
                   data = balanced, REML = TRUE)

# Method 3: Mixed Effects (ML)
mixed_ml <- lmer(Moved_out_of_Bay_w23 ~ within_manual + person_mean + (1 | prim_key),
                 data = balanced, REML = FALSE)

# Method 4: OLS with person dummies (should equal FE)
ols_dummies <- lm(Moved_out_of_Bay_w23 ~ within_manual + factor(prim_key), data = balanced)

# Comparison
test3_results <- data.frame(
  Method = c("Fixed Effects (plm)",
             "Mixed Effects (REML)",
             "Mixed Effects (ML)",
             "OLS with Person Dummies"),
  Coefficient = c(coef(fe_balanced)["lag_Kin_within_60_N"],
                  fixef(mixed_reml)["within_manual"],
                  fixef(mixed_ml)["within_manual"],
                  coef(ols_dummies)["within_manual"]),
  stringsAsFactors = FALSE
)
test3_results$Difference_from_FE <- test3_results$Coefficient - test3_results$Coefficient[1]

styled_table(test3_results, "Test 3: Balanced Panel - All Methods Match")
```

**Result:** All four methods produce identical within-person coefficients.

# Test 4: Multiple Variables

Let's verify the equivalence holds for multiple predictors.

```{r test4-multiple-vars}
# Variables to test
test_vars <- c("lag_Kin_within_60_N", "lag_Nonkin_within_60_N",
               "lag_Owns_Home", "lag_Marr_Cohab")

# Function to compare FE vs Mixed for a given variable
compare_fe_mixed <- function(var, data) {
  # Filter to complete cases for this variable
  data_complete <- data %>%
    filter(!is.na(.data[[var]]), !is.na(Moved_out_of_Bay_w23))

  # Compute decomposition
  data_complete <- data_complete %>%
    group_by(prim_key) %>%
    mutate(
      person_mean = mean(.data[[var]], na.rm = TRUE),
      within_dev = .data[[var]] - person_mean
    ) %>%
    ungroup()

  # FE model
  pdata <- pdata.frame(data_complete, index = c("prim_key", "WAVE"))
  fe_form <- as.formula(paste("Moved_out_of_Bay_w23 ~", var))
  fe_model <- tryCatch(
    plm(fe_form, data = pdata, model = "within"),
    error = function(e) NULL
  )

  if (is.null(fe_model)) return(NULL)

  # Mixed model
  mixed_model <- lmer(Moved_out_of_Bay_w23 ~ within_dev + person_mean + (1 | prim_key),
                      data = data_complete, REML = TRUE)

  data.frame(
    Variable = var,
    N = nrow(data_complete),
    FE_Coef = coef(fe_model)[var],
    Mixed_Within_Coef = fixef(mixed_model)["within_dev"],
    stringsAsFactors = FALSE
  )
}

# Run comparisons
young_data <- df %>% filter(agegroup == 0)
comparisons <- lapply(test_vars, compare_fe_mixed, data = young_data)
comparisons <- do.call(rbind, comparisons[!sapply(comparisons, is.null)])
comparisons$Difference <- comparisons$FE_Coef - comparisons$Mixed_Within_Coef
comparisons$Match <- ifelse(abs(comparisons$Difference) < 1e-6, "Yes", "No")

styled_table(comparisons, "Test 4: Multiple Variables - FE vs Mixed Equivalence")
```

# Summary

```{r summary, results='asis'}
cat("## Validation Results\n\n")

cat("### What We Confirmed\n\n")
cat("1. **The Mundlak approach works**: When decomposition is computed on the same sample,\n")
cat("   within-person coefficients from mixed models EQUAL fixed effects estimates.\n\n")
cat("2. **REML vs ML doesn't matter**: Both produce identical within-person coefficients.\n\n")
cat("3. **The equivalence holds for multiple variables**: Not just one special case.\n\n")

cat("### Implication for Script 10\n\n")
cat("The mixed models in Script 10 use pre-computed decomposition variables.\n")
cat("When filtering by age group, the within coefficients are **close but not identical** to FE.\n\n")
cat("This is acceptable because:\n")
cat("- The differences are small\n")
cat("- The interpretation remains valid (within-person effects)\n")
cat("- For publication-quality results, we could recompute decomposition per model\n\n")

cat("### When to Recompute\n\n")
cat("Recompute within/between decomposition when:\n")
cat("- You need exact FE equivalence for a robustness check\n")
cat("- The sample differs substantially from the full panel\n")
cat("- You want to compare mixed model results directly to FE results\n")
```

# Appendix: How to Recompute Decomposition

For reference, here's how to properly compute within/between decomposition:
code on the **analysis sample**:

```{r appendix-code, eval=FALSE}
# Example: Recompute decomposition for a filtered sample
analysis_data <- df %>%
  filter(agegroup == 0) %>%  # Your filter conditions
  filter(!is.na(outcome_var), !is.na(predictor_var))

# Compute within/between on THIS sample
analysis_data <- analysis_data %>%
  group_by(prim_key) %>%
  mutate(
    between_predictor = mean(predictor_var, na.rm = TRUE),
    within_predictor = predictor_var - between_predictor
  ) %>%
  ungroup()

# Now the mixed model will match FE
model <- lmer(outcome_var ~ within_predictor + between_predictor + (1 | prim_key),
              data = analysis_data)
```
