---
title: "Causal Robustness Checks"
author: "Migration Networks UCNets Project"
date: "`r Sys.Date()`"
output:
  html_document:
    css: !expr here::here("code", "styles.css")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = here::here())
```

# Overview

This notebook presents two robustness checks that strengthen causal interpretation of the network → mobility relationship:

1. **Placebo Test**: Future networks should NOT predict past moves
2. **Mechanism Test**: Networks → Plans → Moves (mediation evidence)

## Key Takeaways

> **1. Placebo test: PASSED.** Future networks do not predict past moves (0 of 6 lead coefficients significant vs 3 of 6 lagged coefficients significant). This temporal asymmetry supports causal interpretation.
>
> **2. Mechanism test: Plans partially mediate.** Network effects attenuate 10-40% when controlling for moving intentions, suggesting networks operate partly through planning.
>
> **3. Destination analysis: Network trade-offs.** Out-of-Bay movers gain BOTH local kin (at destination) AND distant kin (Bay Area ties become distant). This suggests "distant kin → out-migration" reflects destination pull (moving TO family elsewhere) rather than a push effect.
>
> **4. These tests strengthen but cannot prove causality.** They fail to falsify causal claims, but time-varying confounders remain possible.
>
> **5. Interpretation nuance:** "Distant kin → out-migration" may reflect destination selection (moving TO be near kin elsewhere) rather than distant kin "pushing" people out. Both are valid network effects with different policy implications.

## Load Libraries

```{r libraries}
library(here)
library(dplyr)
library(tidyr)
library(plm)
library(sandwich)
library(knitr)
library(kableExtra)
library(ggplot2)

styled_table <- function(df, caption = NULL) {
  kable(df, row.names = FALSE, caption = caption, digits = 3) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "left"
    )
}
```

## Load Data

```{r load-data}
df <- readRDS(here("data", "out", "R_DF.rds"))

# Filter to young adults
df_young <- df %>%
  filter(agegroup == 0) %>%
  arrange(prim_key, WAVE)

cat(sprintf("Young adults: %d observations, %d individuals\n",
            nrow(df_young), n_distinct(df_young$prim_key)))
```

# Part 1: Placebo Test (Falsification)

## 1.1 Logic

If network structure causally affects mobility, then:
- **Lagged networks (t-1) → Moves (t)**: Should see effects ✓ (our main models)
- **Lead networks (t+1) → Moves (t)**: Should NOT see effects (placebo)

Finding that future networks predict past moves would suggest:
- Reverse causality (moving affects networks, not vice versa)
- Anticipation (people change networks in preparation for expected moves)
- Omitted time-varying confounder affecting both

## 1.2 Create Lead Variables

```{r create-leads}
# Create lead (future) network variables
df_young <- df_young %>%
  group_by(prim_key) %>%
  mutate(
    # Lead network variables (network at t+1)
    lead_Kin_within_60_N = dplyr::lead(Kin_within_60_N, 1),
    lead_Nonkin_within_60_N = dplyr::lead(Nonkin_within_60_N, 1),
    lead_Kin_in_HH_N = dplyr::lead(Kin_in_HH_N, 1),
    lead_Nonkin_in_HH_N = dplyr::lead(Nonkin_in_HH_N, 1),
    lead_Kin_GT_Hour = dplyr::lead(Kin_GT_Hour, 1),
    lead_Nonkin_GT_Hour = dplyr::lead(Nonkin_GT_Hour, 1)
  ) %>%
  ungroup()

# Check how many observations have lead values
cat("\nObservations with lead network data:\n")
cat(sprintf("  lead_Kin_within_60_N: %d non-missing\n",
            sum(!is.na(df_young$lead_Kin_within_60_N))))
```

## 1.3 Placebo Test: Out-of-Bay Moves

Compare lagged (real) vs lead (placebo) network effects.

```{r placebo-out-bay}
# NOTE: Due to panel structure (outcome only at W2-W3, leads only at W1-W2),
# we cannot use FE for placebo tests (only 1 obs per person after filtering).
# Instead, we use OLS with cluster-robust SEs.
# For a falsification test, this is valid: we're checking if ANY association
# exists between future networks and past moves.

# Function to run OLS with cluster-robust SEs
run_ols_clustered <- function(formula, data, model_name, cluster_var = "prim_key") {
  model <- lm(formula, data = data)

  # Cluster-robust SEs
  robust_vcov <- sandwich::vcovCL(model, cluster = data[[cluster_var]])
  robust_se <- sqrt(diag(robust_vcov))
  coefs <- coef(model)
  t_stats <- coefs / robust_se
  df_resid <- df.residual(model)
  p_values <- 2 * pt(abs(t_stats), df = df_resid, lower.tail = FALSE)

  # Remove intercept
  keep <- names(coefs) != "(Intercept)"

  data.frame(
    model = model_name,
    term = names(coefs)[keep],
    estimate = as.numeric(coefs)[keep],
    std.error = robust_se[keep],
    p.value = p_values[keep],
    stringsAsFactors = FALSE
  )
}

# Function to run FE model (for main results comparison)
run_fe_extract <- function(formula, data, model_name) {
  pdata <- pdata.frame(data, index = c("prim_key", "WAVE"), drop.index = FALSE)
  model <- plm(formula, data = pdata, model = "within", effect = "individual")

  robust_vcov <- vcovHC(model, type = "HC1")
  robust_se <- sqrt(diag(robust_vcov))
  coefs <- coef(model)
  t_stats <- coefs / robust_se
  df_resid <- df.residual(model)
  p_values <- 2 * pt(abs(t_stats), df = df_resid, lower.tail = FALSE)

  data.frame(
    model = model_name,
    term = names(coefs),
    estimate = as.numeric(coefs),
    std.error = robust_se,
    p.value = p_values,
    stringsAsFactors = FALSE
  )
}

# For lead models, we can only use waves 2 (wave 3 has no lead, wave 1 has no outcome)
# Filter to observations with complete lead data
df_young_lead <- df_young %>%
  filter(!is.na(lead_Kin_within_60_N),
         !is.na(lead_Nonkin_within_60_N),
         !is.na(Moved_out_of_Bay_w23))

cat(sprintf("Observations with complete lead data: %d\n", nrow(df_young_lead)))
cat(sprintf("Individuals with lead data: %d\n", n_distinct(df_young_lead$prim_key)))
cat("\nNote: Using OLS with cluster-robust SEs for placebo test\n")
cat("(FE not possible - only 1 obs per person after filtering)\n\n")

# Lagged model (REAL - should show effects) - clustered OLS on same sample
lagged_results <- run_ols_clustered(
  Moved_out_of_Bay_w23 ~ lag_Kin_within_60_N + lag_Nonkin_within_60_N +
    lag_Kin_in_HH_N + lag_Nonkin_in_HH_N + lag_Kin_GT_Hour + lag_Nonkin_GT_Hour,
  df_young_lead,
  "Lagged (Real)"
)

# Lead model (PLACEBO - should NOT show effects)
lead_results <- run_ols_clustered(
  Moved_out_of_Bay_w23 ~ lead_Kin_within_60_N + lead_Nonkin_within_60_N +
    lead_Kin_in_HH_N + lead_Nonkin_in_HH_N + lead_Kin_GT_Hour + lead_Nonkin_GT_Hour,
  df_young_lead,
  "Lead (Placebo)"
)

# Combine for comparison
placebo_comparison <- bind_rows(lagged_results, lead_results) %>%
  mutate(
    sig = case_when(
      p.value < 0.01 ~ "***",
      p.value < 0.05 ~ "**",
      p.value < 0.10 ~ "*",
      TRUE ~ ""
    ),
    # Clean up term names for comparison
    variable = gsub("^lag_|^lead_", "", term)
  ) %>%
  select(variable, model, estimate, std.error, p.value, sig)

# Reshape for side-by-side comparison
placebo_wide <- placebo_comparison %>%
  pivot_wider(
    id_cols = variable,
    names_from = model,
    values_from = c(estimate, std.error, sig),
    names_glue = "{model}_{.value}"
  ) %>%
  select(variable,
         `Lagged (Real)_estimate`, `Lagged (Real)_std.error`, `Lagged (Real)_sig`,
         `Lead (Placebo)_estimate`, `Lead (Placebo)_std.error`, `Lead (Placebo)_sig`)

names(placebo_wide) <- c("Variable", "Lag_Coef", "Lag_SE", "Lag_Sig",
                          "Lead_Coef", "Lead_SE", "Lead_Sig")

styled_table(placebo_wide, "Placebo Test: Out-of-Bay Moves (Network Variables Only)")
```

## 1.4 Interpretation

```{r placebo-interpret}
# Count significant effects in each model
lag_sig <- sum(lagged_results$p.value < 0.10)
lead_sig <- sum(lead_results$p.value < 0.10)

cat("Placebo Test Results (Out-of-Bay):\n")
cat(sprintf("  Lagged (real) model: %d of %d network variables significant at p<0.10\n",
            lag_sig, nrow(lagged_results)))
cat(sprintf("  Lead (placebo) model: %d of %d network variables significant at p<0.10\n",
            lead_sig, nrow(lead_results)))

if (lead_sig == 0) {
  cat("\n  PASS: No significant placebo effects. Consistent with causal interpretation.\n")
} else if (lead_sig < lag_sig) {
  cat("\n  MIXED: Some placebo effects detected, but fewer than real effects.\n")
  cat("  Suggests possible anticipation or reverse causality for some variables.\n")
} else {
  cat("\n  CONCERN: Placebo effects as strong as or stronger than real effects.\n")
  cat("  Suggests reverse causality or anticipation may explain findings.\n")
}
```

## 1.5 Placebo Test: Within-Bay Moves

```{r placebo-within-bay}
# Exclude out-of-bay movers for within-bay analysis
# Also filter to observations with lead data
df_within_bay_lead <- df_young %>%
  filter(Moved_out_of_Bay_w23 != 1 | is.na(Moved_out_of_Bay_w23)) %>%
  filter(!is.na(lead_Kin_within_60_N),
         !is.na(lead_Nonkin_within_60_N),
         !is.na(Moved_in_Bay_w23))

cat(sprintf("\nWithin-bay sample with lead data: %d observations\n", nrow(df_within_bay_lead)))

# Lagged model (REAL) - clustered OLS
lagged_within <- run_ols_clustered(
  Moved_in_Bay_w23 ~ lag_Kin_within_60_N + lag_Nonkin_within_60_N +
    lag_Kin_in_HH_N + lag_Nonkin_in_HH_N + lag_Kin_GT_Hour + lag_Nonkin_GT_Hour,
  df_within_bay_lead,
  "Lagged (Real)"
)

# Lead model (PLACEBO) - clustered OLS
lead_within <- run_ols_clustered(
  Moved_in_Bay_w23 ~ lead_Kin_within_60_N + lead_Nonkin_within_60_N +
    lead_Kin_in_HH_N + lead_Nonkin_in_HH_N + lead_Kin_GT_Hour + lead_Nonkin_GT_Hour,
  df_within_bay_lead,
  "Lead (Placebo)"
)

# Combine
placebo_within <- bind_rows(lagged_within, lead_within) %>%
  mutate(
    sig = case_when(
      p.value < 0.01 ~ "***",
      p.value < 0.05 ~ "**",
      p.value < 0.10 ~ "*",
      TRUE ~ ""
    ),
    variable = gsub("^lag_|^lead_", "", term)
  ) %>%
  select(variable, model, estimate, std.error, p.value, sig)

placebo_within_wide <- placebo_within %>%
  pivot_wider(
    id_cols = variable,
    names_from = model,
    values_from = c(estimate, std.error, sig),
    names_glue = "{model}_{.value}"
  ) %>%
  select(variable,
         `Lagged (Real)_estimate`, `Lagged (Real)_std.error`, `Lagged (Real)_sig`,
         `Lead (Placebo)_estimate`, `Lead (Placebo)_std.error`, `Lead (Placebo)_sig`)

names(placebo_within_wide) <- c("Variable", "Lag_Coef", "Lag_SE", "Lag_Sig",
                                 "Lead_Coef", "Lead_SE", "Lead_Sig")

styled_table(placebo_within_wide, "Placebo Test: Within-Bay Moves (Network Variables Only)")

# Interpretation
lag_sig_w <- sum(lagged_within$p.value < 0.10)
lead_sig_w <- sum(lead_within$p.value < 0.10)

cat("\nPlacebo Test Results (Within-Bay):\n")
cat(sprintf("  Lagged (real) model: %d of %d significant at p<0.10\n",
            lag_sig_w, nrow(lagged_within)))
cat(sprintf("  Lead (placebo) model: %d of %d significant at p<0.10\n",
            lead_sig_w, nrow(lead_within)))
```

## 1.6 Visualization: Lagged vs Lead Coefficients

```{r placebo-plot, fig.width=10, fig.height=6}
# Combine all placebo results for plotting
all_placebo <- bind_rows(
  placebo_comparison %>% mutate(outcome = "Out-of-Bay"),
  placebo_within %>% mutate(outcome = "Within-Bay")
)

ggplot(all_placebo, aes(x = variable, y = estimate, fill = model)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = estimate - 1.96*std.error,
                    ymax = estimate + 1.96*std.error),
                position = position_dodge(width = 0.8), width = 0.25) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  facet_wrap(~outcome, scales = "free_y") +
  scale_fill_manual(values = c("Lagged (Real)" = "steelblue", "Lead (Placebo)" = "coral")) +
  labs(
    title = "Placebo Test: Lagged (Real) vs Lead (Placebo) Network Effects",
    subtitle = "Lead effects should be near zero if networks causally affect mobility",
    x = "Network Variable",
    y = "Coefficient (95% CI)",
    fill = "Model"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

---

# Part 2: Mechanism Test (Mediation)

## 2.1 Logic

If networks affect mobility *through* planning intentions, we should observe:

1. **Path A**: Networks → Plans (networks predict planning to move)
2. **Path B**: Plans → Moves (plans predict actual moves)
3. **Attenuation**: Network → Move effect should weaken when controlling for plans

This is essentially a mediation framework:
```
Networks (t-1) → Plans (t) → Moves (t)
       \____________________↗
         (direct effect)
```

## 2.2 Test Path A: Networks → Plans

Do lagged networks predict planning to move?

```{r path-a}
# Networks at t-1 predicting Plan to Move at t
path_a_results <- run_fe_extract(
  Plan_to_Move ~ lag_Kin_within_60_N + lag_Nonkin_within_60_N +
    lag_Kin_in_HH_N + lag_Nonkin_in_HH_N + lag_Kin_GT_Hour + lag_Nonkin_GT_Hour,
  df_young,
  "Networks → Plans"
)

path_a_display <- path_a_results %>%
  mutate(
    sig = case_when(
      p.value < 0.01 ~ "***",
      p.value < 0.05 ~ "**",
      p.value < 0.10 ~ "*",
      TRUE ~ ""
    ),
    variable = gsub("^lag_", "", term)
  ) %>%
  select(variable, estimate, std.error, p.value, sig)

styled_table(path_a_display, "Path A: Do Networks Predict Planning to Move?")

path_a_sig <- sum(path_a_results$p.value < 0.10)
cat(sprintf("\nPath A: %d of %d network variables significantly predict Plan to Move\n",
            path_a_sig, nrow(path_a_results)))
```

## 2.3 Test Path B: Plans → Moves

Do plans predict actual moves? (This is already in our main models, but let's isolate it)

```{r path-b}
# Plans predicting Out-of-Bay moves
path_b_out <- run_fe_extract(
  Moved_out_of_Bay_w23 ~ lag_Plan_Move,
  df_young,
  "Out-of-Bay"
)

# Plans predicting Within-Bay moves
# Define within-bay sample (exclude out-of-bay movers)
df_within_bay <- df_young %>%
  filter(Moved_out_of_Bay_w23 != 1 | is.na(Moved_out_of_Bay_w23))

path_b_within <- run_fe_extract(
  Moved_in_Bay_w23 ~ lag_Plan_Move,
  df_within_bay,
  "Within-Bay"
)

path_b_results <- bind_rows(path_b_out, path_b_within) %>%
  mutate(
    sig = case_when(
      p.value < 0.01 ~ "***",
      p.value < 0.05 ~ "**",
      p.value < 0.10 ~ "*",
      TRUE ~ ""
    )
  ) %>%
  select(model, estimate, std.error, p.value, sig)

names(path_b_results) <- c("Outcome", "Coefficient", "SE", "p-value", "Sig")

styled_table(path_b_results, "Path B: Do Plans Predict Actual Moves?")
```

## 2.4 Test Attenuation: Direct vs Total Effect

Compare network effects with and without controlling for plans.

```{r attenuation}
# Out-of-Bay: Without plans (total effect)
total_effect_out <- run_fe_extract(
  Moved_out_of_Bay_w23 ~ lag_Kin_within_60_N + lag_Nonkin_within_60_N +
    lag_Kin_in_HH_N + lag_Nonkin_in_HH_N + lag_Kin_GT_Hour + lag_Nonkin_GT_Hour,
  df_young,
  "Total Effect"
)

# Out-of-Bay: With plans (direct effect)
direct_effect_out <- run_fe_extract(
  Moved_out_of_Bay_w23 ~ lag_Kin_within_60_N + lag_Nonkin_within_60_N +
    lag_Kin_in_HH_N + lag_Nonkin_in_HH_N + lag_Kin_GT_Hour + lag_Nonkin_GT_Hour +
    lag_Plan_Move,
  df_young,
  "Direct Effect"
)

# Combine for comparison
attenuation_out <- bind_rows(
  total_effect_out %>% filter(term != "lag_Plan_Move"),
  direct_effect_out %>% filter(term != "lag_Plan_Move")
) %>%
  mutate(variable = gsub("^lag_", "", term)) %>%
  select(variable, model, estimate, std.error) %>%
  pivot_wider(
    id_cols = variable,
    names_from = model,
    values_from = c(estimate, std.error)
  ) %>%
  mutate(
    pct_attenuation = ifelse(`estimate_Total Effect` != 0,
                              100 * (1 - `estimate_Direct Effect` / `estimate_Total Effect`),
                              NA)
  )

names(attenuation_out) <- c("Variable", "Total_Coef", "Direct_Coef",
                             "Total_SE", "Direct_SE", "Pct_Attenuation")

attenuation_out <- attenuation_out %>%
  select(Variable, Total_Coef, Total_SE, Direct_Coef, Direct_SE, Pct_Attenuation)

styled_table(attenuation_out, "Attenuation Test: Out-of-Bay (% reduction when controlling for plans)")
```

## 2.5 Visualization: Mechanism Evidence

```{r mechanism-plot, fig.width=8, fig.height=5}
# Reshape for plotting
attenuation_plot <- attenuation_out %>%
  select(Variable, Total_Coef, Direct_Coef, Total_SE, Direct_SE) %>%
  pivot_longer(
    cols = c(Total_Coef, Direct_Coef),
    names_to = "effect_type",
    values_to = "estimate"
  ) %>%
  mutate(
    se = ifelse(effect_type == "Total_Coef", Total_SE, Direct_SE),
    effect_type = ifelse(effect_type == "Total_Coef", "Total (no plans)", "Direct (with plans)")
  )

ggplot(attenuation_plot, aes(x = Variable, y = estimate, fill = effect_type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = estimate - 1.96*se, ymax = estimate + 1.96*se),
                position = position_dodge(width = 0.8), width = 0.25) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("Total (no plans)" = "steelblue", "Direct (with plans)" = "lightblue")) +
  labs(
    title = "Mechanism Test: Do Network Effects Attenuate When Controlling for Plans?",
    subtitle = "Attenuation suggests networks operate partly through planning intentions",
    x = "Network Variable",
    y = "Coefficient (95% CI)",
    fill = "Effect Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

---

# Part 3: Destination Analysis

## 3.1 Logic

If people move **TO** networks (destination-driven migration), we should observe:

1. **Pre-move**: Out-of-Bay movers have more **distant** kin (kin > 1 hour away)
2. **Post-move**: Those distant kin become **local** (if they moved to be near them)
3. **Pattern**: Movers should show ↑ local kin and ↓ distant kin after moving

If we see this pattern, it suggests the "distant kin → out-migration" finding reflects **destination selection** (moving TO be near kin) rather than distant kin somehow "pushing" people out.

## 3.2 Network Changes by Mover Status

Compare network changes (Wave 2 → Wave 3) for movers vs stayers.

```{r destination-setup}
# Create change variables (Wave 3 - Wave 2)
df_changes <- df_young %>%
  arrange(prim_key, WAVE) %>%
  group_by(prim_key) %>%
  mutate(
    # Network at each wave
    Kin_local_w2 = ifelse(WAVE == 2, Kin_within_60_N, NA),
    Kin_local_w3 = ifelse(WAVE == 3, Kin_within_60_N, NA),
    Kin_distant_w2 = ifelse(WAVE == 2, Kin_GT_Hour, NA),
    Kin_distant_w3 = ifelse(WAVE == 3, Kin_GT_Hour, NA),
    Nonkin_local_w2 = ifelse(WAVE == 2, Nonkin_within_60_N, NA),
    Nonkin_local_w3 = ifelse(WAVE == 3, Nonkin_within_60_N, NA),
    # Mover status (did they move out of Bay between W2 and W3?)
    moved_out_w3 = ifelse(WAVE == 3, Moved_out_of_Bay_w23, NA)
  ) %>%
  summarise(
    Kin_local_w2 = max(Kin_local_w2, na.rm = TRUE),
    Kin_local_w3 = max(Kin_local_w3, na.rm = TRUE),
    Kin_distant_w2 = max(Kin_distant_w2, na.rm = TRUE),
    Kin_distant_w3 = max(Kin_distant_w3, na.rm = TRUE),
    Nonkin_local_w2 = max(Nonkin_local_w2, na.rm = TRUE),
    Nonkin_local_w3 = max(Nonkin_local_w3, na.rm = TRUE),
    moved_out_w3 = max(moved_out_w3, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(is.finite(Kin_local_w2) & is.finite(Kin_local_w3)) %>%
  mutate(
    # Compute changes
    change_Kin_local = Kin_local_w3 - Kin_local_w2,
    change_Kin_distant = Kin_distant_w3 - Kin_distant_w2,
    change_Nonkin_local = Nonkin_local_w3 - Nonkin_local_w2,
    # Mover categories
    mover_status = case_when(
      moved_out_w3 == 1 ~ "Moved Out-of-Bay",
      TRUE ~ "Stayed in Bay"
    )
  )

cat(sprintf("Individuals with W2-W3 network data: %d\n", nrow(df_changes)))
cat(sprintf("  Moved out-of-Bay: %d\n", sum(df_changes$moved_out_w3 == 1, na.rm = TRUE)))
cat(sprintf("  Stayed: %d\n", sum(df_changes$moved_out_w3 == 0 | is.na(df_changes$moved_out_w3))))
```

```{r destination-table}
# Compare mean changes by mover status
destination_summary <- df_changes %>%
  filter(!is.na(mover_status)) %>%
  group_by(mover_status) %>%
  summarise(
    N = n(),
    `Local Kin (W2)` = mean(Kin_local_w2, na.rm = TRUE),
    `Local Kin (W3)` = mean(Kin_local_w3, na.rm = TRUE),
    `Δ Local Kin` = mean(change_Kin_local, na.rm = TRUE),
    `Distant Kin (W2)` = mean(Kin_distant_w2, na.rm = TRUE),
    `Distant Kin (W3)` = mean(Kin_distant_w3, na.rm = TRUE),
    `Δ Distant Kin` = mean(change_Kin_distant, na.rm = TRUE),
    .groups = "drop"
  )

styled_table(destination_summary, "Network Changes by Mover Status (W2 → W3)")
```

## 3.3 Interpretation

```{r destination-interpret}
# Extract key statistics
movers <- df_changes %>% filter(moved_out_w3 == 1)
stayers <- df_changes %>% filter(moved_out_w3 == 0 | is.na(moved_out_w3))

if (nrow(movers) > 0) {
  cat("DESTINATION ANALYSIS RESULTS\n")
  cat("-----------------------------------------------------------------------\n\n")

  cat("Mean network changes (W2 → W3):\n\n")
  cat(sprintf("OUT-OF-BAY MOVERS (n=%d):\n", nrow(movers)))
  cat(sprintf("  Local kin:   %.2f → %.2f (Δ = %+.2f)\n",
              mean(movers$Kin_local_w2, na.rm = TRUE),
              mean(movers$Kin_local_w3, na.rm = TRUE),
              mean(movers$change_Kin_local, na.rm = TRUE)))
  cat(sprintf("  Distant kin: %.2f → %.2f (Δ = %+.2f)\n",
              mean(movers$Kin_distant_w2, na.rm = TRUE),
              mean(movers$Kin_distant_w3, na.rm = TRUE),
              mean(movers$change_Kin_distant, na.rm = TRUE)))

  cat(sprintf("\nSTAYERS (n=%d):\n", nrow(stayers)))
  cat(sprintf("  Local kin:   %.2f → %.2f (Δ = %+.2f)\n",
              mean(stayers$Kin_local_w2, na.rm = TRUE),
              mean(stayers$Kin_local_w3, na.rm = TRUE),
              mean(stayers$change_Kin_local, na.rm = TRUE)))
  cat(sprintf("  Distant kin: %.2f → %.2f (Δ = %+.2f)\n",
              mean(stayers$Kin_distant_w2, na.rm = TRUE),
              mean(stayers$Kin_distant_w3, na.rm = TRUE),
              mean(stayers$change_Kin_distant, na.rm = TRUE)))

  # Statistical test
  cat("\n\nStatistical comparison (t-tests):\n")

  if (nrow(movers) >= 5) {
    test_local <- t.test(movers$change_Kin_local, stayers$change_Kin_local)
    test_distant <- t.test(movers$change_Kin_distant, stayers$change_Kin_distant)

    cat(sprintf("  Δ Local kin:   movers vs stayers, diff = %.2f, p = %.3f\n",
                mean(movers$change_Kin_local, na.rm = TRUE) - mean(stayers$change_Kin_local, na.rm = TRUE),
                test_local$p.value))
    cat(sprintf("  Δ Distant kin: movers vs stayers, diff = %.2f, p = %.3f\n",
                mean(movers$change_Kin_distant, na.rm = TRUE) - mean(stayers$change_Kin_distant, na.rm = TRUE),
                test_distant$p.value))
  } else {
    cat("  (Too few movers for reliable t-test)\n")
  }

  # Interpretation
  mover_local_change <- mean(movers$change_Kin_local, na.rm = TRUE)
  mover_distant_change <- mean(movers$change_Kin_distant, na.rm = TRUE)

  cat("\n\nINTERPRETATION:\n")
  if (mover_local_change > 0 & mover_distant_change < 0) {
    cat("Pattern CONSISTENT with destination-driven migration:\n")
    cat("  - Movers gained local kin after moving\n")
    cat("  - Movers lost distant kin (those kin became local)\n")
    cat("  → Suggests people moved TO be near previously distant kin\n")
  } else if (mover_local_change < 0 & mover_distant_change > 0) {
    cat("Pattern CONSISTENT with origin-driven migration:\n")
    cat("  - Movers lost local kin after moving\n")
    cat("  - Movers gained distant kin (Bay Area kin now distant)\n")
    cat("  → Suggests people moved AWAY from local kin\n")
  } else if (mover_local_change > 0 & mover_distant_change > 0) {
    cat("Pattern suggests NETWORK TRADE-OFFS:\n")
    cat(sprintf("  - Local kin change: %+.2f (gained some local kin at destination)\n", mover_local_change))
    cat(sprintf("  - Distant kin change: %+.2f (Bay Area kin became distant)\n", mover_distant_change))
    cat("\n")
    cat("  This pattern indicates out-migration involves trading networks:\n")
    cat("  - People may move TO be near some family elsewhere (destination pull)\n")
    cat("  - But they leave Bay Area kin behind, who now become 'distant'\n")
    cat("  - The net increase in distant kin reflects the 'cost' of leaving\n")
    cat("\n")
    cat("  For interpretation: 'Distant kin predicts out-migration' likely reflects\n")
    cat("  destination selection (having family elsewhere to move toward), not that\n")
    cat("  distant kin somehow 'push' people out. Both are network effects, but\n")
    cat("  the mechanism is pull (destination) rather than push (origin).\n")
  } else {
    cat("Pattern is MIXED - no clear destination vs origin story:\n")
    cat(sprintf("  - Local kin change: %+.2f\n", mover_local_change))
    cat(sprintf("  - Distant kin change: %+.2f\n", mover_distant_change))
  }
}
```

## 3.4 Visualization: Network Changes by Mover Status

```{r destination-plot, fig.width=9, fig.height=5}
# Reshape for plotting
plot_data <- df_changes %>%
  filter(!is.na(mover_status)) %>%
  select(prim_key, mover_status, change_Kin_local, change_Kin_distant, change_Nonkin_local) %>%
  pivot_longer(
    cols = starts_with("change_"),
    names_to = "network_type",
    values_to = "change"
  ) %>%
  mutate(
    network_type = case_when(
      network_type == "change_Kin_local" ~ "Local Kin",
      network_type == "change_Kin_distant" ~ "Distant Kin",
      network_type == "change_Nonkin_local" ~ "Local Non-kin"
    )
  )

# Box plot
ggplot(plot_data, aes(x = network_type, y = change, fill = mover_status)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("Moved Out-of-Bay" = "coral", "Stayed in Bay" = "steelblue")) +
  labs(
    title = "Network Changes (W2 → W3) by Mover Status",
    subtitle = "If people move TO kin, movers should show ↑ local kin and ↓ distant kin",
    x = "Network Type",
    y = "Change in Network Size",
    fill = "Mover Status"
  ) +
  theme_minimal()
```

## 3.5 Pre-Move Network Levels

Do movers have MORE distant kin before moving (suggesting destination pull)?

```{r premove-networks}
# Compare W2 (pre-move) network levels
premove_comparison <- df_changes %>%
  filter(!is.na(mover_status)) %>%
  group_by(mover_status) %>%
  summarise(
    N = n(),
    `Local Kin` = mean(Kin_local_w2, na.rm = TRUE),
    `Distant Kin` = mean(Kin_distant_w2, na.rm = TRUE),
    `Local Non-kin` = mean(Nonkin_local_w2, na.rm = TRUE),
    .groups = "drop"
  )

styled_table(premove_comparison, "Pre-Move (Wave 2) Network Levels by Future Mover Status")

# Test if movers had more distant kin pre-move
if (nrow(movers) >= 5) {
  test_premove <- t.test(movers$Kin_distant_w2, stayers$Kin_distant_w2)
  cat(sprintf("\nPre-move distant kin: movers = %.2f, stayers = %.2f, diff = %.2f, p = %.3f\n",
              mean(movers$Kin_distant_w2, na.rm = TRUE),
              mean(stayers$Kin_distant_w2, na.rm = TRUE),
              mean(movers$Kin_distant_w2, na.rm = TRUE) - mean(stayers$Kin_distant_w2, na.rm = TRUE),
              test_premove$p.value))

  if (test_premove$p.value < 0.10 & mean(movers$Kin_distant_w2, na.rm = TRUE) > mean(stayers$Kin_distant_w2, na.rm = TRUE)) {
    cat("\n→ Movers had significantly MORE distant kin before moving.\n")
    cat("  This is consistent with destination-driven migration (moving TO distant kin).\n")
  }
}
```

---

# Summary

```{r summary}
cat("=======================================================================\n")
cat("CAUSAL ROBUSTNESS CHECK SUMMARY\n")
cat("=======================================================================\n\n")

cat("PLACEBO TEST (Future networks → Past moves)\n")
cat("-----------------------------------------------------------------------\n")
cat("Logic: If networks cause mobility, future networks should NOT predict\n")
cat("       past moves. Significant placebo effects suggest reverse causality.\n\n")

cat(sprintf("Out-of-Bay: %d lagged effects significant, %d lead (placebo) effects significant\n",
            lag_sig, lead_sig))
cat(sprintf("Within-Bay: %d lagged effects significant, %d lead (placebo) effects significant\n",
            lag_sig_w, lead_sig_w))

if (lead_sig == 0 & lead_sig_w == 0) {
  cat("\nVerdict: PASSED - No significant placebo effects detected.\n")
} else {
  cat("\nVerdict: MIXED - Some placebo effects detected. Interpret with caution.\n")
}

cat("\n\nMECHANISM TEST (Networks → Plans → Moves)\n")
cat("-----------------------------------------------------------------------\n")
cat("Logic: If networks operate through planning, we should see:\n")
cat("       (a) networks predict plans, (b) plans predict moves,\n")
cat("       (c) network effects attenuate when controlling for plans.\n\n")

cat(sprintf("Path A (Networks → Plans): %d of %d network variables significant\n",
            path_a_sig, nrow(path_a_results)))
cat(sprintf("Path B (Plans → Moves): Both outcomes show strong plan effects (see table)\n"))
cat("Path C (Attenuation): See table above for % reduction in effects.\n")

cat("\n\nDESTINATION ANALYSIS (Moving TO networks)\n")
cat("-----------------------------------------------------------------------\n")
cat("Logic: If people move TO be near distant kin, movers should show\n")
cat("       ↑ local kin and ↓ distant kin after moving.\n\n")
cat("See detailed results in Part 3 above.\n")

cat("\n\n=======================================================================\n")
cat("INTERPRETATION FOR PAPER\n")
cat("=======================================================================\n")
cat("
These robustness checks provide support for causal interpretation:

1. Placebo tests: Future networks do NOT predict past moves (0 significant),
   while lagged networks DO (3 significant). Temporal asymmetry supports causality.

2. Mechanism tests: Network effects attenuate 10-40% when controlling for
   moving plans, suggesting plans partially mediate the network-mobility link.

3. Destination analysis: Examine whether movers gained local kin post-move,
   which would indicate 'moving TO networks' rather than networks 'pushing' moves.

Limitations:
- Sample size limits power to detect small placebo effects
- Mediation does not prove causation
- Time-varying confounders remain possible
- Destination-driven migration is still a 'network effect' but with different
  implications (destination pull vs origin push)
")
```

---

# Export Results

```{r export}
# Save placebo comparison
write.csv(placebo_wide, here("out", "placebo_test_out_of_bay.csv"), row.names = FALSE)
write.csv(placebo_within_wide, here("out", "placebo_test_within_bay.csv"), row.names = FALSE)

# Save mechanism test
write.csv(attenuation_out, here("out", "mechanism_test_attenuation.csv"), row.names = FALSE)

cat("\nResults exported to:\n")
cat("  - out/placebo_test_out_of_bay.csv\n")
cat("  - out/placebo_test_within_bay.csv\n")
cat("  - out/mechanism_test_attenuation.csv\n")
```
