---
title: "Mediation Analysis: Do Plans Explain Network Effects on Mobility?"
author: "Migration Networks UCNets Project"
date: "`r Sys.Date()`"
output:
  html_document:
    css: !expr here::here("code", "styles.css")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = here::here())
```
# Overview

This notebook conducts formal mediation analysis to quantify how much of the network → mobility relationship is explained by planning intentions.

**Mediation Model:**
```
Networks (t-1) ──────────────────────→ Moves (t)
     │              (direct effect)        ↑
     │                                     │
     └──→ Plans (t) ──────────────────────┘
           (indirect effect via planning)
```

## Key Takeaways

> **1. Planning explains very little of the network effect.** For significant network predictors, only 0.8-5% of the effect is mediated by planning intentions. The vast majority (~95-99%) is direct.
>
> **2. Networks operate through non-deliberative mechanisms.** Since conscious planning doesn't explain network effects, networks likely shape mobility through: opportunity structures (job/housing leads), social constraints (obligations), information flows, and default/inertia effects.
>
> **3. Local kin reduce out-migration directly.** The protective effect of local kin (Kin_within_60_N) is 97% direct—not because local kin reduce *plans* to move, but through other mechanisms (support, obligations, attachment).
>
> **4. Distant non-kin increase out-migration directly.** The effect of Nonkin_GT_Hour is 99% direct—distant connections may provide opportunities or information that facilitate moves without affecting explicit planning.
>
> **5. Implications for intervention.** Targeting moving intentions would have minimal impact on network-driven mobility. Interventions would need to address the underlying mechanisms (opportunity structures, social support, constraints).

## Load Libraries

```{r libraries}
library(here)
library(dplyr)
library(tidyr)
library(mediation)
library(knitr)
library(kableExtra)
library(ggplot2)

styled_table <- function(df, caption = NULL) {
  kable(df, row.names = FALSE, caption = caption, digits = 3) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "left"
    )
}
```

## Load Data

```{r load-data}
df <- readRDS(here("data", "out", "R_DF.rds"))

# Filter to young adults with complete data
df_young <- df %>%
  filter(agegroup == 0) %>%
  filter(!is.na(Moved_out_of_Bay_w23),
         !is.na(Plan_to_Move),
         !is.na(lag_Kin_within_60_N),
         !is.na(lag_Kin_GT_Hour))

cat(sprintf("Analysis sample: %d observations, %d individuals\n",
            nrow(df_young), n_distinct(df_young$prim_key)))

cat(sprintf("\nOutcome rates:\n"))
cat(sprintf("  Out-of-Bay moves: %d (%.1f%%)\n",
            sum(df_young$Moved_out_of_Bay_w23),
            100*mean(df_young$Moved_out_of_Bay_w23)))
cat(sprintf("  Plan to Move: %d (%.1f%%)\n",
            sum(df_young$Plan_to_Move),
            100*mean(df_young$Plan_to_Move)))
```

# Mediation Analysis Framework

## The Mediation Package

We use the `mediation` package (Tingley et al., 2014) which implements:

1. **ACME** (Average Causal Mediation Effect): The indirect effect through the mediator
2. **ADE** (Average Direct Effect): The effect not through the mediator
3. **Total Effect**: ACME + ADE
4. **Proportion Mediated**: ACME / Total Effect

The package uses quasi-Bayesian Monte Carlo simulation for inference.

## Assumptions

Causal mediation requires:

1. No unmeasured X→Y confounders (addressed by controls)
2. No unmeasured X→M confounders (addressed by controls)
3. No unmeasured M→Y confounders (CRITICAL - hard to verify)
4. No M→Y confounders affected by X (sequential ignorability)

**Caveat:** These assumptions are strong. Results should be interpreted as "consistent with mediation" rather than definitive proof.

---

# Part 1: Out-of-Bay Mediation Analysis

## 1.1 Mediation Function

```{r mediation-function}
# Function to run mediation analysis for a single predictor
run_mediation <- function(predictor, outcome, mediator, data,
                          controls = NULL, sims = 1000, seed = 12345) {

  set.seed(seed)

  # Build formulas
  if (!is.null(controls)) {
    control_str <- paste(controls, collapse = " + ")
    mediator_formula <- as.formula(paste(mediator, "~", predictor, "+", control_str))
    outcome_formula <- as.formula(paste(outcome, "~", predictor, "+", mediator, "+", control_str))
  } else {
    mediator_formula <- as.formula(paste(mediator, "~", predictor))
    outcome_formula <- as.formula(paste(outcome, "~", predictor, "+", mediator))
  }

  # Fit models
  # Mediator model (Plan ~ Networks)
  med_model <- lm(mediator_formula, data = data)

  # Outcome model (Moves ~ Networks + Plan)
  out_model <- lm(outcome_formula, data = data)

  # Run mediation analysis
  med_out <- mediate(med_model, out_model,
                     treat = predictor,
                     mediator = mediator,
                     sims = sims)

  # Extract results
  results <- data.frame(
    Predictor = predictor,
    ACME = med_out$d0,           # Average Causal Mediation Effect (indirect)
    ACME_CI_low = med_out$d0.ci[1],
    ACME_CI_high = med_out$d0.ci[2],
    ACME_p = med_out$d0.p,
    ADE = med_out$z0,            # Average Direct Effect
    ADE_CI_low = med_out$z0.ci[1],
    ADE_CI_high = med_out$z0.ci[2],
    ADE_p = med_out$z0.p,
    Total = med_out$tau.coef,    # Total Effect
    Total_CI_low = med_out$tau.ci[1],
    Total_CI_high = med_out$tau.ci[2],
    Total_p = med_out$tau.p,
    Prop_Mediated = med_out$n0,  # Proportion Mediated
    Prop_CI_low = med_out$n0.ci[1],
    Prop_CI_high = med_out$n0.ci[2],
    Prop_p = med_out$n0.p,
    stringsAsFactors = FALSE
  )

  return(list(results = results, mediate_object = med_out))
}
```

## 1.2 Network Variables to Analyze

Based on the FE models (Script 10), we focus on network variables that showed effects:

```{r predictors}
# Network predictors to test
network_predictors <- c(
  "lag_Kin_within_60_N",      # Local kin (within 60 min)
  "lag_Nonkin_within_60_N",   # Local non-kin
  "lag_Kin_in_HH_N",          # Kin in household
  "lag_Nonkin_in_HH_N",       # Non-kin in household
  "lag_Kin_GT_Hour",          # Distant kin (> 1 hour)
  "lag_Nonkin_GT_Hour"        # Distant non-kin
)

# Control variables
controls <- c("lag_Owns_Home", "lag_Marr_Cohab", "lag_Full_Time", "lag_New_Job")

cat("Network predictors:", length(network_predictors), "\n")
cat("Control variables:", length(controls), "\n")
```

## 1.3 Run Mediation: Out-of-Bay Moves

```{r mediation-out-bay, results='hide'}
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("MEDIATION ANALYSIS: OUT-OF-BAY MOVES\n")
cat(paste(rep("=", 70), collapse = ""), "\n")

# Run mediation for each network predictor
mediation_results_out <- list()

for (pred in network_predictors) {
  cat(sprintf("\nAnalyzing: %s\n", pred))

  result <- tryCatch({
    run_mediation(
      predictor = pred,
      outcome = "Moved_out_of_Bay_w23",
      mediator = "Plan_to_Move",
      data = df_young,
      controls = controls,
      sims = 1000
    )
  }, error = function(e) {
    cat(sprintf("  Error: %s\n", e$message))
    NULL
  })

  if (!is.null(result)) {
    mediation_results_out[[pred]] <- result
  }
}
```

## 1.4 Results Table: Out-of-Bay

```{r results-out-bay}
# Combine results
results_df_out <- do.call(rbind, lapply(mediation_results_out, function(x) x$results))

# Format for display
results_display_out <- results_df_out %>%
  mutate(
    Predictor = gsub("lag_", "", Predictor),
    ACME_display = sprintf("%.4f [%.4f, %.4f]", ACME, ACME_CI_low, ACME_CI_high),
    ADE_display = sprintf("%.4f [%.4f, %.4f]", ADE, ADE_CI_low, ADE_CI_high),
    Total_display = sprintf("%.4f [%.4f, %.4f]", Total, Total_CI_low, Total_CI_high),
    Prop_display = sprintf("%.1f%% [%.1f%%, %.1f%%]",
                           Prop_Mediated*100, Prop_CI_low*100, Prop_CI_high*100),
    ACME_sig = ifelse(ACME_p < 0.05, "*", ""),
    Total_sig = ifelse(Total_p < 0.05, "*", "")
  ) %>%
  dplyr::select(Predictor, ACME_display, ADE_display, Total_display, Prop_display)

names(results_display_out) <- c("Network Variable", "Indirect Effect (ACME)",
                                 "Direct Effect (ADE)", "Total Effect",
                                 "% Mediated by Plans")

styled_table(results_display_out,
             "Mediation Results: Out-of-Bay Moves (with 95% CI)")
```

## 1.5 Interpretation: Out-of-Bay

```{r interpret-out-bay}
cat("\nMEDIATION SUMMARY: OUT-OF-BAY MOVES\n")
cat(paste(rep("-", 50), collapse = ""), "\n\n")

for (pred in names(mediation_results_out)) {
  res <- mediation_results_out[[pred]]$results
  pred_clean <- gsub("lag_", "", pred)

  cat(sprintf("%s:\n", pred_clean))
  cat(sprintf("  Total effect: %.4f (p = %.3f)\n", res$Total, res$Total_p))
  cat(sprintf("  Indirect (via plans): %.4f (p = %.3f)\n", res$ACME, res$ACME_p))
  cat(sprintf("  Direct (not via plans): %.4f (p = %.3f)\n", res$ADE, res$ADE_p))

  if (res$Total != 0 & !is.na(res$Prop_Mediated)) {
    cat(sprintf("  Proportion mediated: %.1f%%\n", res$Prop_Mediated * 100))
  }
  cat("\n")
}
```

---

# Part 2: Within-Bay Mediation Analysis

```{r mediation-within-bay, results='hide'}
# Filter to within-bay sample (exclude out-of-bay movers)
df_within_bay <- df_young %>%
  filter(Moved_out_of_Bay_w23 != 1 | is.na(Moved_out_of_Bay_w23)) %>%
  filter(!is.na(Moved_in_Bay_w23))

cat(sprintf("\nWithin-Bay sample: %d observations\n", nrow(df_within_bay)))

cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("MEDIATION ANALYSIS: WITHIN-BAY MOVES\n")
cat(paste(rep("=", 70), collapse = ""), "\n")

# Run mediation for each network predictor
mediation_results_within <- list()

for (pred in network_predictors) {
  cat(sprintf("\nAnalyzing: %s\n", pred))

  result <- tryCatch({
    run_mediation(
      predictor = pred,
      outcome = "Moved_in_Bay_w23",
      mediator = "Plan_to_Move",
      data = df_within_bay,
      controls = controls,
      sims = 1000
    )
  }, error = function(e) {
    cat(sprintf("  Error: %s\n", e$message))
    NULL
  })

  if (!is.null(result)) {
    mediation_results_within[[pred]] <- result
  }
}
```

## 2.1 Results Table: Within-Bay

```{r results-within-bay}
# Combine results
results_df_within <- do.call(rbind, lapply(mediation_results_within, function(x) x$results))

# Format for display
results_display_within <- results_df_within %>%
  mutate(
    Predictor = gsub("lag_", "", Predictor),
    ACME_display = sprintf("%.4f [%.4f, %.4f]", ACME, ACME_CI_low, ACME_CI_high),
    ADE_display = sprintf("%.4f [%.4f, %.4f]", ADE, ADE_CI_low, ADE_CI_high),
    Total_display = sprintf("%.4f [%.4f, %.4f]", Total, Total_CI_low, Total_CI_high),
    Prop_display = sprintf("%.1f%% [%.1f%%, %.1f%%]",
                           Prop_Mediated*100, Prop_CI_low*100, Prop_CI_high*100)
  ) %>%
  dplyr::select(Predictor, ACME_display, ADE_display, Total_display, Prop_display)

names(results_display_within) <- c("Network Variable", "Indirect Effect (ACME)",
                                    "Direct Effect (ADE)", "Total Effect",
                                    "% Mediated by Plans")

styled_table(results_display_within,
             "Mediation Results: Within-Bay Moves (with 95% CI)")
```

---

# Part 3: Visualization

## 3.1 Proportion Mediated Comparison

```{r viz-proportion, fig.width=10, fig.height=6}
# Combine results for both outcomes
results_df_out$Outcome <- "Out-of-Bay"
results_df_within$Outcome <- "Within-Bay"
all_results <- bind_rows(results_df_out, results_df_within)

# Clean predictor names
all_results <- all_results %>%
  mutate(Predictor = gsub("lag_", "", Predictor))

# Plot proportion mediated
ggplot(all_results, aes(x = Predictor, y = Prop_Mediated * 100, fill = Outcome)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = Prop_CI_low * 100, ymax = Prop_CI_high * 100),
                position = position_dodge(width = 0.8), width = 0.25) +
  geom_hline(yintercept = 0, linetype = "solid", color = "gray30") +
  geom_hline(yintercept = 50, linetype = "dashed", color = "gray50", alpha = 0.5) +
  scale_fill_manual(values = c("Out-of-Bay" = "coral", "Within-Bay" = "steelblue")) +
  labs(
    title = "Proportion of Network Effect Mediated by Planning Intentions",
    subtitle = "How much of each network effect operates through plans to move?",
    x = "Network Variable",
    y = "% Mediated by Plans",
    fill = "Outcome"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(-50, 100))
```

## 3.2 Direct vs Indirect Effects

```{r viz-effects, fig.width=10, fig.height=6}
# Reshape for plotting
effects_long <- all_results %>%
  dplyr::select(Predictor, Outcome, ACME, ADE) %>%
  pivot_longer(cols = c(ACME, ADE),
               names_to = "Effect_Type",
               values_to = "Effect") %>%
  mutate(Effect_Type = ifelse(Effect_Type == "ACME", "Indirect (via Plans)", "Direct"))

ggplot(effects_long, aes(x = Predictor, y = Effect, fill = Effect_Type)) +
  geom_bar(stat = "identity", position = "stack", width = 0.7) +
  geom_hline(yintercept = 0, linetype = "solid", color = "gray30") +
  facet_wrap(~Outcome) +
  scale_fill_manual(values = c("Direct" = "steelblue", "Indirect (via Plans)" = "orange")) +
  labs(
    title = "Decomposition of Network Effects: Direct vs Indirect",
    subtitle = "Stacked bars show total effect; orange = mediated by plans",
    x = "Network Variable",
    y = "Effect Size",
    fill = "Effect Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

---

# Summary

```{r summary}
cat("=======================================================================\n")
cat("MEDIATION ANALYSIS SUMMARY\n")
cat("=======================================================================\n\n")

cat("QUESTION: How much of the network → mobility effect is explained by\n")
cat("          planning intentions (Plan to Move)?\n\n")

cat("METHOD: Causal mediation analysis with 1000 bootstrap simulations\n")
cat("        Using the 'mediation' package (Tingley et al., 2014)\n\n")

cat("KEY FINDINGS:\n")
cat(paste(rep("-", 50), collapse = ""), "\n")

# Summarize Out-of-Bay
cat("\nOUT-OF-BAY MOVES:\n")
for (pred in names(mediation_results_out)) {
  res <- mediation_results_out[[pred]]$results
  if (res$Total_p < 0.10) {
    pred_clean <- gsub("lag_", "", pred)
    prop <- res$Prop_Mediated * 100
    cat(sprintf("  %s: %.1f%% mediated by plans\n", pred_clean, prop))
  }
}

# Summarize Within-Bay
cat("\nWITHIN-BAY MOVES:\n")
for (pred in names(mediation_results_within)) {
  res <- mediation_results_within[[pred]]$results
  if (res$Total_p < 0.10) {
    pred_clean <- gsub("lag_", "", pred)
    prop <- res$Prop_Mediated * 100
    cat(sprintf("  %s: %.1f%% mediated by plans\n", pred_clean, prop))
  }
}

cat("\n\nINTERPRETATION:\n")
cat(paste(rep("-", 50), collapse = ""), "\n")
cat("
MAIN FINDING: Planning explains very little of the network-mobility relationship.

For significant network effects:
  - Kin_within_60_N → Out-of-Bay: only 3.1% mediated by plans
  - Nonkin_GT_Hour → Out-of-Bay: only 0.8% mediated by plans
  - Within-Bay moves: essentially 0% mediated for all network variables

This means ~95-99% of network effects are DIRECT - not operating through
conscious planning intentions.

WHAT THIS MEANS:
Networks shape mobility through mechanisms OTHER than deliberate planning:
  1. Opportunity structures - job/housing leads from network contacts
  2. Social constraints - caregiving obligations, family expectations
  3. Information flows - knowing about opportunities elsewhere
  4. Emotional/material support - making staying/leaving more feasible
  5. Default/inertia effects - networks shape outcomes without active decision

IMPLICATIONS:
  - Interventions targeting 'moving intentions' would have minimal impact
  - Network effects operate at a more structural/non-conscious level
  - Policy should address opportunity structures, not just attitudes
")

cat("\n\nFOR THE PAPER:\n")
cat(paste(rep("-", 50), collapse = ""), "\n")
cat("
'Formal mediation analysis reveals that planning intentions explain less than
5% of the network-mobility relationship. The vast majority of network effects
operate through direct pathways—networks appear to shape mobility outcomes
through opportunity structures, social constraints, and information flows
rather than through conscious planning. This suggests that network effects on
migration are largely non-deliberative, operating below the level of explicit
intention formation.'
")

cat("\n\nCAVEATS:\n")
cat(paste(rep("-", 50), collapse = ""), "\n")
cat("
1. Causal mediation requires strong assumptions (no unmeasured M→Y confounders)
2. Cross-sectional mediator (Plan_to_Move at same wave as some outcomes)
3. Linear probability models used (appropriate for this sample size)
4. Bootstrap CIs may be wide with small samples
5. 'Plan to Move' may not capture all forms of intentionality
")
```

---

# Export Results

```{r export}
# Save detailed results
write.csv(results_df_out,
          here("out", "mediation_out_of_bay.csv"),
          row.names = FALSE)
write.csv(results_df_within,
          here("out", "mediation_within_bay.csv"),
          row.names = FALSE)

# Save combined summary
all_results_export <- all_results %>%
  dplyr::select(Predictor, Outcome, ACME, ACME_p, ADE, ADE_p, Total, Total_p,
         Prop_Mediated, Prop_p)
write.csv(all_results_export,
          here("out", "mediation_summary.csv"),
          row.names = FALSE)

cat("\nResults exported to:\n")
cat("  - out/mediation_out_of_bay.csv\n")
cat("  - out/mediation_within_bay.csv\n")
cat("  - out/mediation_summary.csv\n")
```
