---
title: "Data Preparation Checks"
author: "Migration Networks UCNets Project"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)
```

# Overview

This notebook runs diagnostic checks on the prepared dataset.
Run this after `run_all_prep.R` to verify data quality.

## Load Libraries and Data

```{r load-data}
library(here)
library(dplyr)

# Load the prepared dataset
df <- readRDS(here("data", "out", "R_DF.rds"))

cat(sprintf("Loaded dataset: %d observations, %d variables\n", nrow(df), ncol(df)))
```

# Basic Data Checks

## WAVE Distribution

```{r wave-distribution}
cat("\nWAVE distribution:\n")
print(table(df$WAVE))
```

## WAVE by Moved_w23 Cross-tabulation

Wave 1 has more NAs because `Moved_w23` captures moves between waves 2-3, so Wave 1 values were imputed from `a19c_onlyw1` (time at residence). The total per wave (~1159) is consistent, indicating a balanced panel. NAs represent respondents who didn't answer or dropped out.

```{r wave-moved-crosstab}
cat("\nWAVE by Moved_w23 cross-tabulation:\n")
print(table(df$WAVE, df$Moved_w23, useNA = "ifany"))
```

# Multicollinearity Checks

## Correlation Matrix

No concerning multicollinearity: all correlations are below 0.4. Notable moderate correlations (|r| > 0.25) include homeownership with kin in household (0.34) and less moving (-0.33), and a trade-off between kin/nonkin in household (-0.28). Patterns are substantively sensible - homeownership associates with stability, while renting/roommates associates with mobility.

```{r multicollinearity-check}
cor_vars <- c("Marr_Cohab", "Full_Time", "LEX_FinanProb", "LT_BA", "LE_NewJob",
              "lag_New_School", "lag_Moved_Town", "Plan_to_Move", "lag_Kin_in_5_N",
              "lag_Nonkin_in_5_N", "lag_Nonkin_in_HH_N", "lag_Kin_in_HH_N", "lag_Owns_Home")

cat("\nCorrelation matrix for multicollinearity check:\n")
cor_matrix <- df %>%
  select(all_of(cor_vars)) %>%
  cor(use = "pairwise.complete.obs")
print(round(cor_matrix, 3))
```

## VIF Check

Nothing concerning here either.
```{r vif-check}
vif_model <- lm(Moved_w23 ~ Marr_Cohab + Full_Time + LEX_FinanProb + LT_BA +
                  LE_NewJob + lag_New_School + lag_Moved_Town + Plan_to_Move +
                  lag_Kin_in_5_N + lag_Nonkin_in_5_N + lag_Nonkin_in_HH_N +
                  lag_Kin_in_HH_N + lag_Owns_Home,
                data = df)

cat("\nVIF values:\n")
if (requireNamespace("car", quietly = TRUE)) {
  print(car::vif(vif_model))
} else {
  cat("Install 'car' package for VIF calculation: install.packages('car')\n")
}
```

# Exploratory Analysis

## Out-of-Bay Movers

Examine young people who moved out of Bay in multiple waves. Some individuals appear to have moved out of the Bay in both W2 and W3. Per CSF memo (2/27/2024), this is likely due to: (1) mis-remembering and reporting the same move twice, (2) sequential moves where respondents interpreted the question geographically (e.g., moved NY→FL and answered "Out of California"), or (3) rare return-then-leave cases (only ~2 cases). Not considered a critical data issue.

**Detailed case review identified 5 cases where W3 responses imply local moves** (e.g., "apartment too loud", "housemate issues", "house was sold", "leaving parent's house", "roommate issues") rather than out-of-Bay moves. These are corrected in `01_create_long_variables.Rmd` by recoding `Moved_out_of_Bay_w23 = 0` at Wave 3 for: 5000000000027, 6000000003997, 7000000003067, 8000000000852, 8000000000885. The remaining double-movers appear to be legitimate sequential long-distance moves (e.g., Bay→Hawaii→New Zealand, Bay→China→grad school).

```{r out-of-bay-analysis}
out_of_bay_young <- df %>%
  filter(Moved_out_of_Bay_w23 == 1) %>%
  filter(WAVE != 1) %>%
  filter(agegroup == 0) %>%
  group_by(prim_key) %>%
  mutate(prim_key_dup = n() - 1) %>%
  ungroup()

cat("\nDuplicate prim_key counts among young out-of-Bay movers:\n")
print(table(out_of_bay_young$prim_key_dup))

# Keep only those with duplicates
out_of_bay_duplicates <- out_of_bay_young %>%
  filter(prim_key_dup == 1) %>%
  select(prim_key, Moved_out_of_Bay_w23)

cat("\nYoung individuals who moved out of Bay in both Wave 2 and 3:\n")
print(out_of_bay_duplicates)
```

# Summary

```{r summary}
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("DATA CHECKS COMPLETE\n")
cat(paste(rep("=", 70), collapse = ""), "\n")
```
