---
title: "Methods"
format: docx
bibliography: "Fixed and Hybrid Effects Models.bib"
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(here)
library(flextable)
library(dplyr)
library(tidyr)
library(officer)

set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 10,
  padding.top = 1,
  padding.bottom = 1
)

# Load data early so inline references work
data_path <- here("data", "out", "R_DF.rds")
if (!file.exists(data_path)) data_path <- here("data", "R_DF.rds")
df_raw <- readRDS(data_path)

# Sample counts (used in inline text before desc-stats chunk)
n_total_obs <- nrow(df_raw)
n_total_resp <- n_distinct(df_raw$prim_key)
n_young_resp <- n_distinct(df_raw$prim_key[df_raw$agegroup == 0])
n_old_resp <- n_distinct(df_raw$prim_key[df_raw$agegroup == 1])
```

# Data and Methods

## The UCNets Study

We use data from the UC Berkeley Social Networks Study (UCNets). UCNets surveyed 1,159 respondents living in the San Francisco Bay Area in two age cohorts (young adults aged 21–30 and older adults aged 50–70) three times between 2015 and 2018 using a combination of in-person and online methods. The study was designed to draw egocentric maps of respondents' personal networks and to collect detailed information about their social connections, socioeconomic circumstances, and health.

The panel structure yields three observations per respondent and up to two inter-wave transitions during which residential mobility can occur. The full analysis sample includes `r n_total_resp` respondents (`r n_young_resp` young, `r n_old_resp` older) contributing `r format(n_total_obs, big.mark=",")` person-waves. The unit of analysis is the person-wave. We focus on moves that occurred within the study period. All analyses are unweighted. UCNets provides demographic survey weights, but our panel models rely on within-person variation that is not improved by cross-sectional reweighting. Weighted robustness checks are reported in Appendix E.

## Moving-Related Questions in the Survey

At each wave, respondents were asked whether they had moved since the last interview and, if so, whether within the Bay Area or out of the region. Respondents who moved were also asked an open-ended question about why they moved ("Why did you move?"), which we coded into categories (described in the Results). Separately, respondents were asked whether they planned to move before the next interview, providing a prospective measure of mobility intentions that can be compared to actual subsequent behavior.

## Outcome Variables

We examine three sets of outcomes. First, for the analysis of who moves and where, we use binary indicators of *planning to move* at a given wave, *moving out of the Bay Area* between consecutive waves, and *moving within the Bay Area* between waves. For within-Bay models, we exclude respondents who moved out of the region so that the comparison is between local movers and stayers only.

Second, for the analysis of self-reported reasons for moving, we use the coded open-ended responses. The categorization process involved multiple steps: a large language model scanned responses to suggest preliminary categories, which were then refined through manual analysis and review of existing literature. Multiple coders, including the authors, categorized the data, allowing for multiple categories per response. The authors conducted manual reviews to ensure accuracy. The resulting categories include career (own or partner's job/school), partner/relationship, financial, housing, new environment, family/friends, conflict/forced, and other.

Third, for the analysis of moving's network consequences, we use network count variables as outcomes: local non-kin (within 60 minutes), local kin, distant non-kin (more than 60 minutes away), and distant kin. These are the same network measures described below, but treated as dependent variables rather than predictors.

## Predictor Variables

Network data were collected through name eliciting questions in which respondents identified their social ties and reported the geographic proximity of each alter. We construct counts of kin and non-kin ties at three distance thresholds: *household* (living in the respondent's home; non-kin household ties primarily reflect roommates among young adults), *local* (within 60 minutes), and *distant* (more than 60 minutes away). This yields six network variables (kin and non-kin at each of three distances), all measured as counts that vary across waves.[^1]

[^1]: These count variables are right-skewed, but we retain untransformed counts for interpretability; results are substantively unchanged with winsorization.

Time-varying socioeconomic controls include homeownership, full-time employment, education (BA or higher vs. less than BA), and marital/cohabitation status. Life events are binary indicators for experiences since the prior interview: starting a new job, enrolling in school, experiencing financial problems, and having a child. We also include a lagged indicator of whether a close contact moved away.

Time-invariant controls include gender, race/ethnicity (Black, Asian, Latino/a, and other race, with white as the reference), and age. Survey mode (online vs. in-person) is included as a between-person control and a within-person control.[^2]

[^2]: Survey mode can independently affect how many ties respondents name. Because a substantial share of non-movers also switched modes across waves, there is sufficient independent variation to include mode as a time-varying control in the fixed effects models, helping to isolate the effect of moving from the effect of changing survey mode.

## Descriptive Overview

```{r desc-stats, include=FALSE}
# Compute observation-level descriptive stats for narrative text
# (df_raw and sample counts already loaded in setup chunk)

pf <- function(x) sprintf("%.1f", 100 * mean(x, na.rm = TRUE))

# Observation-level subgroups
y_obs <- df_raw %>% filter(agegroup == 0)
o_obs <- df_raw %>% filter(agegroup == 1)

y_within  <- pf(y_obs$Moved_in_Bay_w23)
y_out     <- pf(y_obs$Moved_out_of_Bay_w23)
o_within  <- pf(o_obs$Moved_in_Bay_w23)
o_out     <- pf(o_obs$Moved_out_of_Bay_w23)
y_plan    <- pf(y_obs$Plan_to_Move)
o_plan    <- pf(o_obs$Plan_to_Move)

y_stay_obs  <- y_obs %>% filter(Moved_out_of_Bay_w23 == 0 & Moved_in_Bay_w23 == 0)
y_local_obs <- y_obs %>% filter(Moved_in_Bay_w23 == 1)
y_dist_obs  <- y_obs %>% filter(Moved_out_of_Bay_w23 == 1)

y_dist_ba    <- pf(y_dist_obs$BA_or_greater)
y_stay_ba    <- pf(y_stay_obs$BA_or_greater)
y_dist_own   <- pf(y_dist_obs$Owns_home)
y_stay_own   <- pf(y_stay_obs$Owns_home)
y_dist_marr  <- pf(y_dist_obs$Marr_Cohab)
y_stay_marr  <- pf(y_stay_obs$Marr_Cohab)
y_local_ba   <- pf(y_local_obs$BA_or_greater)
y_local_ft   <- pf(y_local_obs$Full_Time)
y_stay_ft    <- pf(y_stay_obs$Full_Time)
y_local_marr <- pf(y_local_obs$Marr_Cohab)

o_stay_obs  <- o_obs %>% filter(Moved_out_of_Bay_w23 == 0 & Moved_in_Bay_w23 == 0)
o_local_obs <- o_obs %>% filter(Moved_in_Bay_w23 == 1)
o_dist_obs  <- o_obs %>% filter(Moved_out_of_Bay_w23 == 1)

o_dist_marr <- pf(o_dist_obs$Marr_Cohab)
o_local_blk <- pf(o_local_obs$black_race)
o_local_own <- pf(o_local_obs$Owns_home)
o_stay_own  <- pf(o_stay_obs$Owns_home)

y_newjob  <- pf(y_obs$LE_NewJob)
y_finprob <- pf(y_obs$LE_FinanProb)
```

Table 1 presents sample characteristics by mobility status and age group at the observation level (person-waves). Mobility is far more common among young adults: `r y_within`% of young adult person-waves involve a within-Bay move and `r y_out`% an out-of-Bay move, compared to `r o_within`% and `r o_out`% among older adults. Planning to move is similarly concentrated among the young (`r y_plan`% vs. `r o_plan`%).

Among young adults, out-of-Bay movers are more educated (`r y_dist_ba`% BA or higher vs. `r y_stay_ba`% of stayers), less likely to be homeowners (`r y_dist_own`% vs. `r y_stay_own`%), and more likely to be married or cohabiting (`r y_dist_marr`% vs. `r y_stay_marr`%). Within-Bay movers show a similar education gradient (`r y_local_ba`% BA+), higher rates of full-time employment (`r y_local_ft`% vs. `r y_stay_ft`%), and higher marriage/cohabitation rates (`r y_local_marr`% vs. `r y_stay_marr`%). Among older adults, the small number of movers limits inference, but out-of-Bay movers are disproportionately married (`r o_dist_marr`%) while within-Bay movers are disproportionately Black (`r o_local_blk`%) and less likely to be homeowners (`r o_local_own`% vs. `r o_stay_own`%).

Life events also differ by mover status. New jobs are reported in `r y_newjob`% of young adult person-waves and financial problems in `r y_finprob`%. Network variables show that young adults have more non-kin household ties (reflecting roommates) and more distant non-kin, while older adults have more local kin. A respondent-level version of this table using baseline characteristics is in Appendix Table S27.

```{r tbl-descriptive, eval=TRUE}
# ---------------------------------------------------------------------------
# Descriptive Table: Observation-level characteristics by move status
# ---------------------------------------------------------------------------
  df_desc <- df_raw %>%
    mutate(
      move_status = case_when(
        Moved_out_of_Bay_w23 == 1 ~ "Out of Bay",
        Moved_in_Bay_w23 == 1     ~ "Within Bay",
        TRUE                      ~ "Non-Mover"
      )
    )

  fmt_pct_t <- function(x) sprintf("%.1f%%", 100 * mean(x, na.rm = TRUE))
  fmt_mean_sd_t <- function(x) sprintf("%.1f (%.1f)", mean(x, na.rm = TRUE), sd(x, na.rm = TRUE))

  build_col <- function(data) {
    n <- nrow(data)
    c(
      fmt_pct_t(data$female), fmt_mean_sd_t(data$age),
      fmt_pct_t(data$black_race), fmt_pct_t(data$asian_race), fmt_pct_t(data$latino_eth),
      fmt_pct_t(data$BA_or_greater), fmt_pct_t(data$Owns_home),
      fmt_pct_t(data$Full_Time), fmt_pct_t(data$Marr_Cohab),
      fmt_pct_t(data$LE_NewJob), fmt_pct_t(data$LE_NewSchool),
      fmt_pct_t(data$LE_FinanProb), fmt_pct_t(data$LEX_Birth),
      fmt_mean_sd_t(data$Kin_in_HH_N), fmt_mean_sd_t(data$Kin_within_60_N),
      fmt_mean_sd_t(data$Kin_GT_Hour_N), fmt_mean_sd_t(data$Nonkin_in_HH_N),
      fmt_mean_sd_t(data$Nonkin_within_60_N), fmt_mean_sd_t(data$Nonkin_GT_Hour_N),
      as.character(n)
    )
  }

  var_names <- c(
    "Female (%)", "Age (mean, SD)",
    "Black (%)", "Asian (%)", "Latino/a (%)",
    "BA or Higher (%)", "Homeowner (%)",
    "Full-Time Employed (%)", "Married/Cohabiting (%)",
    "New Job (%)", "New School (%)",
    "Financial Problems (%)", "Birth of Child (%)",
    "Kin in Household", "Kin within 60 min",
    "Kin >60 min", "Non-kin in Household",
    "Non-kin within 60 min", "Non-kin >60 min",
    "N (person-waves)"
  )

  section_labels <- c("Demographics", "Socioeconomic Status",
                       "Life Events", "Network Variables (mean, SD)")

  y_non_t   <- build_col(df_desc %>% filter(agegroup == 0, move_status == "Non-Mover"))
  y_local_t <- build_col(df_desc %>% filter(agegroup == 0, move_status == "Within Bay"))
  y_dist_t  <- build_col(df_desc %>% filter(agegroup == 0, move_status == "Out of Bay"))
  o_non_t   <- build_col(df_desc %>% filter(agegroup == 1, move_status == "Non-Mover"))
  o_local_t <- build_col(df_desc %>% filter(agegroup == 1, move_status == "Within Bay"))
  o_dist_t  <- build_col(df_desc %>% filter(agegroup == 1, move_status == "Out of Bay"))

  rows <- list()
  for (i in seq_along(var_names)) {
    if (i == 1)  rows[[length(rows) + 1]] <- c(section_labels[1], rep("", 6))
    if (i == 6)  rows[[length(rows) + 1]] <- c(section_labels[2], rep("", 6))
    if (i == 10) rows[[length(rows) + 1]] <- c(section_labels[3], rep("", 6))
    if (i == 14) rows[[length(rows) + 1]] <- c(section_labels[4], rep("", 6))
    rows[[length(rows) + 1]] <- c(
      paste0("  ", var_names[i]),
      y_non_t[i], y_local_t[i], y_dist_t[i],
      o_non_t[i], o_local_t[i], o_dist_t[i]
    )
  }

  tbl <- as.data.frame(do.call(rbind, rows), stringsAsFactors = FALSE)
  names(tbl) <- c("Variable", "YN", "YL", "YD", "ON", "OL", "OD")

  sec_rows <- which(tbl$Variable %in% section_labels)
  n_row <- which(grepl("N \\(person", tbl$Variable))

  ft <- flextable(tbl)
  ft <- set_header_labels(ft, Variable = "",
    YN = "Non-Mover", YL = "Local", YD = "Distant",
    ON = "Non-Mover", OL = "Local", OD = "Distant")
  ft <- add_header_row(ft,
    values = c("", "Young Adults (21-30)", "Older Adults (50-70)"),
    colwidths = c(1, 3, 3), top = TRUE)
  ft <- theme_booktabs(ft)
  ft <- bold(ft, i = sec_rows, j = 1, part = "body")
  if (length(n_row) > 0) {
    ft <- hline(ft, i = n_row - 1, part = "body",
                border = fp_border(width = 0.5))
  }
  ft <- fontsize(ft, size = 9, part = "all")
  ft <- font(ft, fontname = "Times New Roman", part = "all")
  ft <- align(ft, j = 1, align = "left", part = "body")
  ft <- align(ft, j = -1, align = "center", part = "body")
  ft <- align(ft, align = "center", part = "header")
  ft <- bold(ft, part = "header")
  ft <- set_caption(ft, caption = "Table 1. Sample Characteristics by Mobility Status and Age Group (Observation-Level)")
  ft <- add_footer_lines(ft, values = paste(
    "Observation-level: each person-wave is one row.",
    "Local = moved within the Bay Area; Distant = moved out of the Bay Area.",
    "Percentages for binary variables; mean (SD) for continuous variables."
  ))
  ft <- fontsize(ft, size = 9, part = "footer")
  ft <- font(ft, fontname = "Times New Roman", part = "footer")
  ft <- italic(ft, part = "footer")
  ft <- autofit(ft)
  ft
```

## Analytical Strategy

The panel includes three waves, yielding up to two transitions per individual. We pool transitions across waves with wave fixed effects. Full-sample analyses (planning, mobility) are run in parallel for young and older adults. Within-mover analyses (reasons for moving, network consequences) focus on young adults in the main text, with older adult results in an appendix, given the low mobility rate among older adults.

For the analysis of planning and mobility, we estimate cross-sectional linear probability models (LPM) that pool observations across all waves, with wave fixed effects and cluster-robust standard errors by individual to account for repeated observations. Although logistic regression is conventional for binary outcomes, LPM offers more straightforward interpretation of coefficients as changes in probability [@brumback_adjusting_2010]. These models address the question of what characteristics distinguish planners and movers from non-planners and stayers.

To estimate how moving affects networks, we use two panel approaches. The first is an individual fixed effects model with network counts as dependent variables:

$$
N_{it} = \alpha_i + \beta_1 \text{MoveOut}_{it} + \beta_2 \text{MoveLocal}_{it} + \boldsymbol{\gamma} \mathbf{X}_{it} + \epsilon_{it}
$$

where $N_{it}$ is the network count (e.g., local non-kin) for individual $i$ at wave $t$, $\alpha_i$ is an individual fixed effect absorbing all time-invariant characteristics, $\text{MoveOut}_{it}$ and $\text{MoveLocal}_{it}$ indicate whether the respondent moved out of or within the Bay Area since the prior wave, and $\mathbf{X}_{it}$ is a vector of time-varying controls (homeownership, employment, education, marital status, life events, plan to move, and survey mode). Because respondents reported whether they moved *since the last interview* before their current network was measured, the move necessarily precedes the network measurement within each wave, establishing temporal ordering. These models include survey mode (web vs. in-person) as a time-varying control, because respondents who moved out of the Bay Area could no longer participate in person and nearly all switched to web administration; controlling for mode ensures that apparent network changes after an out-of-Bay move do not simply reflect differences in how networks are reported online. We use heteroskedasticity-consistent (HC1) robust standard errors to account for potential heteroskedasticity in the error terms.

To understand lasting effects on respondent networks, we use a first-difference model in the following form:

$$
\Delta N_{i} = N_{i,w3} - N_{i,w2} = \beta_1 \text{MoveOut}_{i,w2} + \beta_2 \text{MoveLocal}_{i,w2} + \boldsymbol{\gamma} \mathbf{X}_{i,w2} + \epsilon_{i}
$$

where the dependent variable is the change in network ties from wave 2 to wave 3, and the predictors are whether the respondent moved between waves 1 and 2 plus wave-2 controls. The model includes all respondents observed at both waves 2 and 3; the coefficients of interest identify the effect by comparing respondents who moved between waves 1 and 2 against non-movers in terms of their subsequent network change from wave 2 to wave 3. By placing a full wave between the move and the network outcome, this design captures whether network disruptions observed immediately after a move persist, stabilize, or reverse over time. We use cluster-robust standard errors by individual.

## References
